---
title: "Figure 3k–m: DREADD Activation of Lepr/Glp1r Neurons"
author: "Dylan Rausch"
output: html_notebook
---

Chemogenetic activation of Lepr/Glp1r neurons using an intersectional
Cre/Dre-dependent DREADD approach (Trh-Cre;Glp1r-Dre). Within-subject
crossover design (CNO vs saline). Three paradigms:

- **3k**: Dark cycle food intake (first 3h)
- **3l**: Refeeding after overnight fast
- **3m**: Ghrelin-induced hyperphagia

## Setup

```{r setup, message=FALSE}
library(tidyverse)
library(lme4)
library(car)
library(emmeans)
library(cowplot)

# -- Paths ---------------------------------------------------------------------
out_dir <- here::here("output_figures/paper/extended_figures/ed_fig3_dreadd")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```

## Load & Prepare Data

```{r load-dreadd}
dreadd_data <- readxl::read_xlsx(here::here("paper/food_intake_analysis/testing.xlsx"), sheet = 3) %>% filter(Time != 0)
```

## Model: Dark Cycle Intake (Figure 3k)

```{r model-darkcycle}
# Log-time model for cumulative intake
mod_darkcycle <- dreadd_data %>%
  filter(Exp == "DarkCycle") %>%
  pivot_longer(starts_with("FEN")) %>%
  mutate(value = log1p(value),
         log_time = log(Time)) %>%
  lmer(value ~ Trt * log_time + (1 | name), data = .)

car::Anova(mod_darkcycle, type = "II")

# Factor-time model for timepoint-specific post-hoc
mod_darkcycle_fac <- dreadd_data %>%
  filter(Exp == "DarkCycle") %>%
  pivot_longer(starts_with("FEN")) %>%
  mutate(value = log1p(value), time = factor(Time)) %>%
  lmer(value ~ Trt * time + (1 | name), data = .)

emmeans(mod_darkcycle_fac, ~ Trt * time) %>%
  pairs(by = "time") %>%
  data.frame()
```

## Model: Fasting-Refeeding (Figure 3l)

```{r model-refeed}
mod_refeed <- dreadd_data %>%
  filter(Exp == "Refeed") %>%
  pivot_longer(starts_with("FEN")) %>%
  mutate(value = log1p(value),
         log_time = log(Time)) %>%
  lmer(value ~ Trt * log_time + (1 | name), data = .)

car::Anova(mod_refeed, type = "II")

mod_refeed_fac <- dreadd_data %>%
  filter(Exp == "Refeed") %>%
  pivot_longer(starts_with("FEN")) %>%
  mutate(value = log1p(value), time = factor(Time)) %>%
  lmer(value ~ Trt * time + (1 | name), data = .)

emmeans(mod_refeed_fac, ~ Trt * time) %>%
  pairs(by = "time") %>%
  data.frame()
```

## Model: Ghrelin-Induced Feeding (Figure 3m)

```{r model-ghrelin}
mod_ghrelin <- dreadd_data %>%
  filter(Exp == "Ghrelin") %>%
  pivot_longer(starts_with("FEN")) %>%
  mutate(value = log1p(value),
         log_time = log(Time)) %>%
  lmer(value ~ Trt * log_time + (1 | name), data = .)

car::Anova(mod_ghrelin, type = "II")

mod_ghrelin_fac <- dreadd_data %>%
  filter(Exp == "Ghrelin") %>%
  pivot_longer(starts_with("FEN")) %>%
  mutate(value = log1p(value), time = factor(Time)) %>%
  lmer(value ~ Trt * time + (1 | name), data = .)

emmeans(mod_ghrelin_fac, ~ Trt * time) %>%
  pairs(by = "time") %>%
  data.frame()
```

## Summary

All three paradigms are modeled as `log1p(cumulative_intake) ~ Treatment * log(Time)`
with mouse as random intercept (within-subject crossover). Log-time is used because
cumulative intake follows a logarithmic time course. The primary result is the main
effect of CNO from `car::Anova()`. Post-hoc pairwise contrasts at each timepoint
use a factor-time model and are provided via `emmeans`.
