---
title: "Figure 4c–d: Leptin Requirement for Feeding Suppression"
author: "Dylan Rausch"
output: html_notebook
---

Tests whether leptin's ability to suppress feeding requires intact LepR signaling
in Glp1r neurons. Compares WT vs Glp1rLepr-KO mice treated with leptin or saline.

- **4c**: Fasting-refeeding with leptin/saline pretreatment
- **4d**: Ghrelin-induced feeding with leptin pretreatment

## Setup

```{r setup, message=FALSE}
library(tidyverse)
library(lme4)
library(car)
library(emmeans)
library(cowplot)
library(patchwork)

# -- Paths ---------------------------------------------------------------------
feeding_file <- here::here("paper/food_intake_analysis/Agrp_Glp1R LepRflox Feeding_TP.DBR 05.2023.xlsx")
out_dir <- here::here("output_figures/paper/extended_figures/ed_fig4_pheno")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# -- Sample metadata -----------------------------------------------------------
mice_data <- readxl::read_xlsx(feeding_file, sheet = 1) %>%
  janitor::clean_names() %>%
  dplyr::select(mouse, strain, genotype, sex)

# -- Color palette -------------------------------------------------------------
cols <- c("#56C1AF", "#155C5B")
```

## Figure 4c — Leptin + Fasting-Refeeding

### Data Processing

```{r process-lep-refeed}
lep_refeed_data <- readxl::read_xlsx(feeding_file, sheet = 5) %>%
  janitor::clean_names() %>%
  arrange(mouse) %>%
  mutate(datetime = as.POSIXct(paste(date, format(time, "%H:%M:%S")),
                               format = "%Y-%m-%d %H:%M:%S")) %>%
  # Leptin/saline experiments (exclude plain refeed experiments)
  filter(!grepl("refeed", treatment)) %>%
  mutate(trt = ifelse(grepl("Lep", treatment), "Lep", "Sal")) %>%
  dplyr::select(-time) %>%
  group_by(mouse, trt) %>%
  mutate(
    expstage = case_when(
      row_number() < min(which(is.na(food))) ~ "baseline",
      row_number() == min(which(is.na(food))) ~ "start-fast",
      row_number() > min(which(is.na(food))) ~ "refeed"
    ),
    start_refeed = min(datetime[expstage == "refeed"]),
    timefromrefeed = as.numeric(difftime(datetime, start_refeed, units = "hours")),
    diet = ifelse(grepl("chow", treatment), "chow", "dio")
  ) %>%
  inner_join(mice_data, by = "mouse") %>%
  group_by(mouse, expstage, strain, trt) %>%
  mutate(
    prev_intake = lag(food) - food,
    cum_intake = cumsum(replace(prev_intake, is.na(prev_intake), 0)),
    geno = ifelse(grepl("KO", genotype), "KO", "WT")
  ) %>%
  group_by(mouse, trt) %>%
  mutate(
    min = min(datetime),
    timefromstart = difftime(datetime, min, units = "hours")
  ) %>%
  filter(!grepl("Agrp", strain)) %>%
  # Baseline food intake: mean hourly intake over 20h pre-fast
  dplyr::select(timefromrefeed, trt, mouse, bw, food, datetime, expstage,
                geno, sex, prev_intake, cum_intake)
```

### Combined Model (Primary Analysis)

**IMPORTANT:** The primary test is the `trt:geno` interaction in the combined
model, NOT separate within-genotype p-values. This avoids the "difference
between significant and not significant" fallacy (Nieuwenhuis et al., 2011).

```{r model-4c-combined}
# Log-time model: cumulative intake follows logarithmic time course
lep_refeed_mod_data <- lep_refeed_data %>%
  filter(timefromrefeed > 0, timefromrefeed < 26) %>%
  mutate(log_time = log(timefromrefeed))

mod_4c <- lmer(log1p(cum_intake) ~ log_time * trt * geno + sex + (1 | mouse),
               data = lep_refeed_mod_data)

# Primary result: trt:geno interaction
car::Anova(mod_4c, type = "II")
```

### Post-hoc Contrasts

```{r posthoc-4c}
# Timepoint-specific contrasts use factor-time model (for plot annotations)
mod_4c_fac <- lep_refeed_data %>%
  filter(timefromrefeed > 0, timefromrefeed < 26) %>%
  mutate(time = factor(timefromrefeed)) %>%
  lmer(log1p(cum_intake) ~ time * trt * geno + sex + (1 | mouse), data = .)

pval_lep_refeed <- emmeans(mod_4c_fac, ~ trt * time * geno) %>%
  pairs(by = c("time", "geno"), adjust = "none") %>%
  data.frame() %>%
  filter(time != 0) %>%
  mutate(adj = p.adjust(p.value, method = "BH"))

pval_lep_refeed
```

### Plot

```{r plot-4c-bars, fig.width=6, fig.height=3}
# Bar plots at 4h and 24h timepoints
ref4 <- lep_refeed_data %>%
  mutate(group = interaction(geno, trt)) %>%
  filter(timefromrefeed == 4) %>%
  ggplot(aes(x = group, y = cum_intake, fill = trt)) +
  stat_summary(geom = "bar", width = 0.8, color = "black") +
  geom_line(aes(group = mouse), color = "grey90") +
  geom_point() +
  facet_wrap(vars(geno), scales = "free_x", strip.position = "bottom") +
  labs(x = "Group", y = "Cumulative Intake (g)") +
  cowplot::theme_cowplot(font_size = 8) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        panel.spacing = unit(0, units = "line"),
        legend.position = "none") +
  scale_fill_manual(values = cols) +
  lims(y = c(0, 5))

ref24 <- lep_refeed_data %>%
  mutate(group = interaction(geno, trt)) %>%
  filter(timefromrefeed == 24) %>%
  ggplot(aes(x = group, y = cum_intake, fill = trt)) +
  stat_summary(geom = "bar", width = 0.8, color = "black") +
  geom_line(aes(group = mouse), color = "grey90") +
  geom_point() +
  facet_wrap(vars(geno), scales = "free_x", strip.position = "bottom") +
  labs(x = "Group", y = "Cumulative Intake (g)") +
  cowplot::theme_cowplot(font_size = 8) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        strip.background = element_blank(),
        panel.spacing = unit(0, units = "line"),
        legend.position = "none") +
  scale_fill_manual(values = cols) +
  lims(y = c(0, 10))

ref4 + ref24
ggsave(file.path(out_dir, "lep_refeed_bars.pdf"), h = 3, w = 6)
```

---

## Figure 4d — Leptin + Ghrelin (1 mg/kg Leptin Pretreatment)

```{r model-4d}
lep_ghrelin_data <- readxl::read_xlsx(feeding_file, sheet = 7) %>%
  janitor::clean_names() %>%
  filter(grepl("Ghrelin \\(1mg", treatment), diet == "Chow") %>%
  inner_join(mice_data, by = "mouse") %>%
  arrange(mouse, treatment, date) %>%
  mutate(
    food = as.numeric(food),
    geno = ifelse(grepl("KO", genotype), "KO", "WT"),
    trt = ifelse(grepl("Leptin", treatment), "Lep+Ghrel", "Sal+Ghrel")
  ) %>%
  group_by(mouse, trt) %>%
  mutate(
    timefromfood = as.numeric(difftime(date, min(date), units = "hours")),
    prev_intake = lag(food) - food,
    cum_intake = cumsum(replace(prev_intake, is.na(prev_intake), 0))
  ) %>%
  filter(timefromfood == 24)

fit_lep_ghrelin <- lmer(log1p(cum_intake) ~ geno * trt + (1 | mouse), data = lep_ghrelin_data)
car::Anova(fit_lep_ghrelin)

pval_lg <- emmeans(fit_lep_ghrelin, pairwise ~ trt | geno)$contrasts %>%
  as.data.frame() %>%
  mutate(adj = p.adjust(p.value, method = "BH"), label = case_when(adj < 0.001 ~ "***", adj < 0.01 ~ "**", adj < 0.05 ~ "*", TRUE ~ ""))
pval_lg
```

### Plot

```{r plot-4d, fig.width=4, fig.height=3.5}
cols_trt <- c("Sal+Ghrel" = "#56C1AF", "Lep+Ghrel" = "#155C5B")

bt_lg <- lep_ghrelin_data %>%
  group_by(geno, trt) %>%
  summarise(m = mean(cum_intake, na.rm = TRUE),
            se = sd(cum_intake, na.rm = TRUE) / sqrt(n()), .groups = "drop") %>%
  group_by(geno) %>%
  summarise(top = max(m + se), .groups = "drop")

sig_lg <- pval_lg %>%
  left_join(bt_lg, by = "geno") %>%
  mutate(y_brack = top * 1.12, y_text = top * 1.28) %>%
  filter(label != "")

p_lep_ghrelin <- lep_ghrelin_data %>%
  ggplot(aes(x = trt, y = cum_intake, fill = trt)) +
  stat_summary(geom = "bar", width = 0.8, color = "black", fun = mean) +
  stat_summary(geom = "errorbar", width = 0.1, fun.data = mean_se) +
  geom_line(aes(group = mouse), color = "grey80", linewidth = 0.4) +
  geom_point(size = 0.8) +
  facet_wrap(vars(geno), scales = "free_x", strip.position = "bottom") +
  geom_segment(data = sig_lg,
               aes(x = 1, xend = 2, y = y_brack, yend = y_brack),
               inherit.aes = FALSE, linewidth = 0.3) +
  geom_segment(data = sig_lg,
               aes(x = 1, xend = 1, y = y_brack * 0.95, yend = y_brack),
               inherit.aes = FALSE, linewidth = 0.3) +
  geom_segment(data = sig_lg,
               aes(x = 2, xend = 2, y = y_brack * 0.95, yend = y_brack),
               inherit.aes = FALSE, linewidth = 0.3) +
  geom_text(data = sig_lg, aes(x = 1.5, y = y_text, label = label),
            inherit.aes = FALSE, size = 2.5) +
  scale_fill_manual(values = cols_trt) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.35))) +
  cowplot::theme_cowplot(font_size = 8) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        strip.background = element_blank(), panel.spacing = unit(0, "line"),
        legend.position = "none") +
  labs(x = NULL, y = "24 h food intake (g)")

ggsave(file.path(out_dir, "lep_ghrelin.pdf"), plot = p_lep_ghrelin, h = 3.5, w = 4)
```

## Summary

**Figure 4c** uses a combined model: `log1p(cum_intake) ~ log(time) * trt * geno + sex + (1|mouse)`.
The primary result is the `trt:geno` interaction, which tests whether leptin's
effect on refeeding differs between WT and KO. Post-hoc contrasts within each
genotype are conditional on this interaction.

**Figure 4d** uses: `log1p(cum_intake) ~ geno * treatment + (1|mouse)` at the
24h timepoint. The `geno:treatment` interaction tests whether leptin's ability to
blunt ghrelin-induced feeding differs by genotype.
