---
title: "Figure 5b–c: HFD Onset — Body Weight, Food Intake, and Circadian Patterns"
author: "Dylan Rausch"
output: html_notebook
---

Glp1rLepr-KO vs control mice switched from chow to HFD. Tests whether leptin
signaling in Glp1r neurons is required for the normal attenuation of hyperphagia
during the onset of obesity.

- **5b (top)**: Body weight trajectory
- **5b (bottom)**: Cumulative food intake
- **5c**: Daily food intake and dark-cycle vs light-cycle breakdown

## Setup

```{r setup, message=FALSE}
library(tidyverse)
library(lme4)
library(car)
library(emmeans)
library(cowplot)
library(lubridate)

# -- Paths ---------------------------------------------------------------------
feeding_file <- here::here("paper/food_intake_analysis/Agrp_Glp1R LepRflox Feeding_TP.DBR 05.2023.xlsx")
out_dir <- here::here("output_figures/paper/extended_figures/ed_fig5_hfd")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# -- Sample groups -------------------------------------------------------------
sample_groups_df <- data.frame(
  SampleID = c(
    "Q0944", "Q0946", "Q0948", "R0084", "R0085", "R0086", "R0089", "R0090", "R0093",
    "Q0941", "Q0942", "Q0943", "Q0945", "Q0949", "Q0950", "R0087", "R0088", "R0091", "R0092"
  ),
  Group = c(rep("Glp1R_LepR_KO", 9), rep("Control", 10))
)

cols <- c("#EBB395", "#97BCDA")
```

## Process Metabolic Cage Data

```{r process-data}
processed_data <- readxl::read_xlsx(feeding_file, sheet = 3,
  col_types = c("date", "date", "text", "numeric", "numeric", "text", "text", "text")) %>%
  dplyr::select(-Notes, -`...8`) %>%
  inner_join(sample_groups_df, by = c("Mouse" = "SampleID")) %>%
  janitor::clean_names() %>%
  transmute(
    date, food, body_weight,
    diet = factor(diet),
    mouse = factor(mouse),
    group = factor(group),
    hour = format(time, format = "%H:%M:%S", tz = "UTC"),
    datetime = as.POSIXct(paste(date, hour), format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
  ) %>%
  arrange(mouse) %>%
  group_by(mouse, diet, group) %>%
  mutate(
    lag = lag(food),
    intake = lag - food,
    time_diff = as.numeric(datetime - lag(datetime)),
    kcal = ifelse(grepl("Chow", diet), intake * 2.89, intake * 5.24),
    day_diff = as.numeric(datetime - lag(datetime)),
    date_adjusted = as.Date(datetime - hours(12))
  )
```

## Figure 5b (top) — Body Weight

### Model

```{r model-bw}
bw_data <- processed_data %>%
  group_by(mouse, date, group, diet) %>%
  summarise(bw = max(body_weight, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.infinite(bw)) %>%
  group_by(mouse) %>%
  mutate(
    norm = mean(bw[date < "2025-01-06"], na.rm = TRUE),
    time = factor((date - min(date)) / (60 * 60 * 24)),
    plot_x = as.numeric(as.character(time))
  )

# LMM: body weight ~ genotype * time, normalized to baseline, HFD only
model_bw <- lmer(log(bw) ~ group * factor(time) + norm + (1 | mouse),
                 data = bw_data %>% filter(!grepl("Chow", diet), plot_x < 14))
car::Anova(model_bw)

pval_bw <- emmeans(model_bw, pairwise ~ group | time) %$%
  contrasts %>%
  data.frame() %>%
  mutate(adj = p.adjust(p.value, method = "BH"))

pval_bw
```

### Plot

```{r plot-bw, fig.width=6, fig.height=4}
bw_plot <- bw_data %>%
  ggplot(aes(x = plot_x, y = bw / norm, color = group, fill = group)) +
  stat_summary(geom = "point", size = 2) +
  stat_summary(geom = "line", linewidth = 1) +
  stat_summary(geom = "ribbon", alpha = 0.5, linewidth = 0) +
  geom_text(
    data = pval_bw %>%
      inner_join(bw_data %>% filter(group == "Glp1R_LepR_KO") %>%
                   group_by(time, plot_x) %>% summarise(mean = mean(bw / norm), .groups = "drop"),
                 by = c("time")) %>%
      mutate(
        label = case_when(adj < 0.001 ~ "***", adj < 0.01 ~ "**", adj < 0.05 ~ "*", TRUE ~ ""),
        y.position = mean + 0.05
      ),
    aes(x = plot_x, y = y.position, label = label), inherit.aes = FALSE
  ) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) +
  scale_x_continuous(breaks = seq(0, 60, by = 7)) +
  cowplot::theme_cowplot(font_size = 8) +
  labs(x = "Days on HFD", y = "Body Weight (normalized)")

bw_plot
ggsave(file.path(out_dir, "bw_hfd.pdf"), h = 4, w = 6)
```

## Figure 5b (bottom) — Cumulative Food Intake

### Model

```{r model-cum-intake}
hfd_full_intake <- processed_data %>%
  filter(!is.na(intake)) %>%
  group_by(mouse, date_adjusted, diet, group) %>%
  reframe(sum = sum(kcal), tot_time = sum(time_diff)) %>%
  group_by(mouse) %>%
  mutate(
    cum_int = cumsum(sum),
    norm = mean(sum[date_adjusted < "2025-01-06"], na.rm = TRUE),
    time = factor(1:n()),
    day_num = as.numeric(as.character(time)),
    log_day = log(day_num)
  )

# Log-time model for cumulative intake (consistent with other cumulative models)
model_cum <- lmer(log(cum_int) ~ group * log_day + norm + (1 | mouse),
                  data = hfd_full_intake %>% filter(!grepl("Chow", diet)))
car::Anova(model_cum, type = "II")

# Factor-time model for timepoint-specific contrasts (plot annotations)
model_cum_fac <- lmer(log(cum_int) ~ group * factor(time) + norm + (1 | mouse),
                      data = hfd_full_intake %>% filter(!grepl("Chow", diet)))

pval_cum <- emmeans(model_cum_fac, pairwise ~ group | time) %$%
  contrasts %>%
  data.frame() %>%
  mutate(adj = p.adjust(p.value, method = "BH"))

pval_cum
```

### Plot

```{r plot-cum-intake, fig.width=6, fig.height=4}
cum_kcal <- hfd_full_intake %>%
  ggplot(aes(x = as.numeric(as.character(time)), y = cum_int,
             color = group, fill = group)) +
  stat_summary(geom = "point", size = 2) +
  stat_summary(geom = "line", linewidth = 1) +
  stat_summary(geom = "ribbon", alpha = 0.5, linewidth = 0) +
  geom_text(
    data = pval_cum %>%
      inner_join(hfd_full_intake %>% filter(group == "Glp1R_LepR_KO") %>%
                   group_by(time) %>% summarise(mean = mean(cum_int), .groups = "drop"),
                 by = "time") %>%
      mutate(
        label = case_when(adj < 0.001 ~ "***", adj < 0.01 ~ "**", adj < 0.05 ~ "*", TRUE ~ ""),
        y.position = mean + 10
      ),
    aes(x = as.numeric(as.character(time)), y = y.position, label = label),
    inherit.aes = FALSE
  ) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) +
  cowplot::theme_cowplot(font_size = 8) +
  labs(x = "Day", y = "Cumulative Intake (kcal)")

cum_kcal
ggsave(file.path(out_dir, "cum_kcal.pdf"), h = 4, w = 6)
```

## Figure 5c — Daily Food Intake

### Model

```{r model-daily-intake}
hfd_model_data <- hfd_full_intake %>%
  mutate(time = factor(time))

model_daily <- lmer(log(sum) ~ group * time + norm + (1 | mouse),
                    data = hfd_model_data %>% filter(!grepl("Chow", diet)))
car::Anova(model_daily, type = "II")

pval_daily <- emmeans(model_daily, pairwise ~ group | time) %$%
  contrasts %>%
  data.frame() %>%
  mutate(adj = p.adjust(p.value, method = "BH"))

pval_daily
```

### Plot

```{r plot-daily-intake, fig.width=6, fig.height=4}
hfd_full_intake %>%
  ggplot(aes(x = as.numeric(as.character(time)), y = sum,
             color = group, fill = group)) +
  stat_summary(geom = "point", size = 3) +
  stat_summary(geom = "line", linewidth = 1) +
  stat_summary(geom = "ribbon", alpha = 0.5) +
  geom_vline(xintercept = 3, lty = 2, linewidth = 1, color = "grey50") +
  geom_text(
    data = pval_daily %>%
      inner_join(hfd_full_intake %>% filter(group == "Glp1R_LepR_KO") %>%
                   group_by(time) %>% summarise(mean = mean(sum), .groups = "drop"),
                 by = "time") %>%
      mutate(
        label = case_when(adj < 0.001 ~ "***", adj < 0.01 ~ "**", adj < 0.05 ~ "*", TRUE ~ ""),
        y.position = mean + 2
      ),
    aes(x = as.numeric(as.character(time)), y = y.position, label = label),
    inherit.aes = FALSE
  ) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) +
  cowplot::theme_cowplot(font_size = 8) +
  labs(x = "Day", y = "Food Intake (kcal)")

ggsave(file.path(out_dir, "kcal_dio.pdf"), h = 4)
```

## Figure 5c — Dark Cycle vs Light Cycle

### Models

```{r model-circadian}
circ_data <- processed_data %>%
  group_by(mouse, date_adjusted, group) %>%
  mutate(pct = kcal / sum(kcal), diet, n = length(unique(diet))) %>%
  group_by(date_adjusted) %>%
  mutate(time = day(date_adjusted) - 2) %>%
  ungroup() %>%
  mutate(night = ifelse(hour < 12, "DC", "LC")) %>%
  group_by(mouse) %>%
  mutate(
    night_kcal = mean(kcal[diet == "5LOD Chow" & night == "DC"], na.rm = TRUE),
    day_kcal = mean(kcal[diet == "5LOD Chow" & night == "LC"], na.rm = TRUE),
    norm_kcal = mean(kcal[diet == "5LOD Chow"], na.rm = TRUE)
  ) %>%
  ungroup()

# --- Dark cycle model (Type II) ---
night_intake <- circ_data %>%
  filter(night == "DC") %>%
  mutate(time = as.factor(time))

mod_night <- lmer(log(kcal) ~ night_kcal + group * time + (1 | mouse),
                  data = night_intake %>% filter(!grepl("Chow", diet)))
car::Anova(mod_night, type = "II")

pval_night <- emmeans(mod_night, pairwise ~ group | time) %$%
  contrasts %>%
  data.frame() %>%
  mutate(adj = p.adjust(p.value, method = "BH"))

pval_night

# --- Light cycle model (Type II for consistency) ---
day_intake <- circ_data %>%
  filter(night == "LC") %>%
  mutate(time = as.factor(time))

mod_day <- lmer(log(kcal) ~ day_kcal + group * time + (1 | mouse),
                data = day_intake %>% filter(!grepl("Chow", diet)))
car::Anova(mod_day, type = "II")

pval_day <- emmeans(mod_day, pairwise ~ group | time) %$%
  contrasts %>%
  data.frame() %>%
  mutate(adj = p.adjust(p.value, method = "BH"))

pval_day
```

### Plot

```{r plot-circadian, fig.width=6, fig.height=6}
circ_data %>%
  filter(time < 10) %>%
  mutate(time = factor(time),
         norm_kcal = ifelse(night == "DC", night_kcal, day_kcal)) %>%
  ggplot(aes(x = time, y = kcal / norm_kcal, lty = group, color = group, fill = group)) +
  stat_summary(geom = "point", size = 3) +
  stat_summary(geom = "line", linewidth = 1) +
  stat_summary(geom = "ribbon", alpha = 0.5, linewidth = 0) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) +
  cowplot::theme_cowplot(font_size = 8) +
  geom_vline(xintercept = 3, lty = 2, linewidth = 1, color = "grey50") +
  facet_wrap(~ night, ncol = 1, scales = "free_y") +
  labs(x = "Day", y = "Food Intake (normalized to baseline)")

ggsave(file.path(out_dir, "circ_intake.pdf"), h = 6, w = 6)
```

## Extended Data Figure 6b — Individual mouse daily intake traces

```{r individual-traces, fig.width=6, fig.height=4}
hfd_full_intake %>%
  filter(!grepl("Chow", diet)) %>%
  ggplot(aes(x = as.numeric(as.character(time)), y = sum, group = mouse, color = group)) +
  geom_line(alpha = 0.4, linewidth = 0.5) +
  stat_summary(aes(group = group), geom = "line", linewidth = 1.5, fun = mean) +
  geom_vline(xintercept = 3, lty = 2, linewidth = 0.6, color = "grey50") +
  scale_color_manual(values = cols) +
  cowplot::theme_cowplot(font_size = 8) +
  labs(x = "Day on HFD", y = "Daily food intake (kcal)", color = NULL)

ggsave(file.path(out_dir, "ed6b_individual_traces.pdf"), h = 4, w = 6)
```

## Extended Data — Cumulative intake individual traces

```{r individual-cum-traces, fig.width=6, fig.height=4}
hfd_full_intake %>%
  filter(!grepl("Chow", diet)) %>%
  ggplot(aes(x = as.numeric(as.character(time)), y = cum_int, group = mouse, color = group)) +
  geom_line(alpha = 0.4, linewidth = 0.5) +
  stat_summary(aes(group = group), geom = "line", linewidth = 1.5, fun = mean) +
  scale_color_manual(values = cols) +
  cowplot::theme_cowplot(font_size = 8) +
  labs(x = "Day on HFD", y = "Cumulative food intake (kcal)", color = NULL)

ggsave(file.path(out_dir, "ed6b_individual_cum_traces.pdf"), h = 4, w = 6)
```

## Summary

All models use `group * factor(time) + baseline_covariate + (1|mouse)`.
The primary result for each panel is the `group:time` interaction from
`car::Anova()`, testing whether genotype differences in food intake or body
weight diverge over time. Post-hoc contrasts at each timepoint are BH-corrected.

**Note:** Both dark-cycle and light-cycle models now use Type II tests for
consistency. The baseline covariate is period-specific (dark-cycle baseline for
the dark model, light-cycle baseline for the light model).
