---
title: "pSTAT3 Spatial Transcriptomics Analysis"
subtitle: "Integration of QuPath IHC classification with 10x Xenium data"
output:
  html_notebook:
    toc: true
    toc_float: true
date: "Originally run June 2024; cleaned March 2026"
---

## Overview

This notebook integrates pSTAT3 IHC classification (from QuPath) with matched 10x Xenium
spatial transcriptomics data. The workflow:

1. Loads Xenium data for 16 tissue regions (2 slides × 8 regions)
2. Maps QuPath-classified pSTAT3+ cells onto Xenium cell segmentations via spatial join
3. Identifies Lepr+ neurons using Gaussian mixture models
4. Subclusters Lepr+ neurons and analyzes pSTAT3 enrichment by diet condition

**QuPath classification details:** See `pstat3_methods.docx` for full documentation of the
cell detection, ANN-MLP classifier (ann_stat3.json), and coordinate transformation pipeline.

## Setup

```{r setup, message=FALSE}
library(Seurat)
library(sf)
library(mclust)
library(tidyverse)
library(ggalluvial)
library(ggsignif)

out_dir <- here::here("output_figures/paper/figure1/")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

options(future.globals.maxSize = 8 * 1024^3)  # 8 GiB for merging large spatial objects

xenium_dir <- here::here("sc_preprocessing/xenium_qupath")
```

## 1. Load Xenium data and map pSTAT3+ cells

For each tissue region, we:

- Load the Xenium output as a Seurat object
- Read the transformed pSTAT3+ cell centroids (classified by QuPath, transformed to Xenium coordinates)
- Perform a spatial join to identify which Xenium cells overlap with pSTAT3+ centroids
- Label overlapping cells as "pos" in the Seurat metadata

```{r load-xenium-737, message=FALSE}
# Slide 21737: regions 1-8, sections 1-8
val <- 1:8
s.list.737 <- lapply(val, function(x) {
  obj <- LoadXenium(paste0(xenium_dir, "/xenium_output/",
                           "output-XETG00161__0021737__Region_", x, "__20240529__110948/"),
                    segmentations = "cell")

  # Read transformed pSTAT3+ centroids (from QuPath ann_stat3 classifier, affine-transformed)
  pos_cells <- data.table::fread(
    paste0(xenium_dir, "/stat3_transformed_ann/renamed_", x, "_cells.csv.gz")
  ) %>%
    data.frame() %>%
    dplyr::rename(x = x_centroid, y = y_centroid)

  # Convert Xenium segmentation polygons to sf
  sf_object <- st_as_sf(obj@images$fov@boundaries$segmentations)
  sf_object[["cell_id.xenium"]] <- colnames(obj)

  # Convert pSTAT3+ centroids to sf points and spatially join with Xenium polygons
  sf_points <- st_as_sf(pos_cells, coords = c("x", "y"), crs = st_crs(sf_object))
  joined <- st_join(sf_object, sf_points)

  # Label Xenium cells that contain a pSTAT3+ centroid
  cellstolab <- joined$cell_id.xenium[!is.na(joined$cell_id)]
  obj[["stat3"]] <- ifelse(colnames(obj) %in% cellstolab, "pos", "neg")
  obj[["section"]] <- rep(x, ncol(obj))
  obj[["slide"]] <- "737"
  return(obj)
})
```

```{r load-xenium-767, message=FALSE}
# Slide 21767: regions 1-8, sections 9-16
val <- 9:16
s.list.767 <- lapply(val, function(x) {
  y <- x - 8
  obj <- LoadXenium(paste0(xenium_dir, "/xenium_output/",
                           "output-XETG00161__0021767__Region_", y, "__20240529__110948/"),
                    segmentations = "cell")

  pos_cells <- data.table::fread(
    paste0(xenium_dir, "/stat3_transformed_ann/renamed_", x, "_cells.csv.gz")
  ) %>%
    data.frame() %>%
    dplyr::rename(x = x_centroid, y = y_centroid)

  sf_object <- st_as_sf(obj@images$fov@boundaries$segmentations)
  sf_object[["cell_id.xenium"]] <- colnames(obj)
  sf_points <- st_as_sf(pos_cells, coords = c("x", "y"), crs = st_crs(sf_object))
  joined <- st_join(sf_object, sf_points)

  cellstolab <- joined$cell_id.xenium[!is.na(joined$cell_id)]
  obj[["stat3"]] <- ifelse(colnames(obj) %in% cellstolab, "pos", "neg")
  obj[["section"]] <- rep(x, ncol(obj))
  obj[["slide"]] <- "767"
  return(obj)
})
```

## 2. Merge and annotate metadata

```{r merge-and-annotate}
seur <- merge(s.list.767[[1]], c(s.list.767[-1], s.list.737))

# Assign diet groups: sections 1-4 & 9-12 = chow; 5-8 & 13-16 = DIO (high-fat diet)
seur[["diet"]] <- ifelse(seur$section %in% c(1:4, 9:12), "chow", "dio")

# Assign animal IDs
seur[["animal"]] <- case_when(
  seur$section %in% 1:3        ~ "animal1",
  seur$section %in% c(4, 12)   ~ "animal2",
  seur$section %in% 5:7        ~ "animal3",
  seur$section %in% c(8, 16)   ~ "animal4",
  seur$section %in% 9:11       ~ "animal5",
  seur$section %in% 13:15      ~ "animal6"
)

seur <- JoinLayers(seur)
seur <- NormalizeData(seur, normalization.method = "CLR", margin = 2)
```

## 3. Cell type classification using Gaussian mixture models

Classify cells as positive/negative for marker genes using 2-component GMMs
on CLR-normalized expression (non-zero cells only).

```{r gmm-classification-functions}
classify_and_assign_cells <- function(seurat_obj, gene_list, assay = "RNA",
                                       n_components = 2, plot = TRUE) {
  seurat_obj <- NormalizeData(seurat_obj, normalization.method = "CLR",
                               margin = 2, assay = assay)

  classify_gene <- function(gene) {
    expr <- GetAssayData(seurat_obj, layer = "data", assay = assay)[gene, ]
    det <- expr[expr > 0]

    gmm <- Mclust(det, G = n_components)
    high_component <- which.max(gmm$parameters$mean)
    is_positive <- names(det)[gmm$classification == high_component]
    vec <- names(expr) %in% is_positive

    if (plot) {
      p <- ggplot(data.frame(expression = det), aes(x = expression)) +
        geom_histogram(aes(y = after_stat(density)), bins = 50, color = "black") +
        geom_density(color = "red") +
        ggtitle(paste("Expression Distribution for", gene)) +
        theme_minimal()
      for (i in 1:n_components) {
        p <- p + stat_function(
          fun = dnorm,
          args = list(mean = gmm$parameters$mean[i],
                      sd = sqrt(gmm$parameters$variance$sigmasq[i])),
          aes(color = factor(i))
        )
      }
      print(p)
    }
    return(vec)
  }

  classification_df <- as.data.frame(sapply(gene_list, classify_gene))
  rownames(classification_df) <- colnames(seurat_obj)

  # Filter for Lepr+ cells, then assign subtypes
  lepr_pos_df <- classification_df %>% filter(Lepr == TRUE) %>% select(-Lepr)

  assign_cell_type <- function(row) {
    positive_genes <- names(row)[row]
    if (length(positive_genes) == 0) return("Negative")
    else if (length(positive_genes) == 1) return(positive_genes)
    else return("Doublet")
  }

  cell_assignments <- apply(lepr_pos_df, 1, assign_cell_type)
  result_df <- data.frame(cell_type = cell_assignments, stringsAsFactors = FALSE) %>%
    rownames_to_column("cell_id")

  return(list(full_classification = classification_df,
              lepr_positive_assignments = result_df))
}
```

```{r run-gmm-classification}
genes_of_interest <- c("Agrp", "Glp1r", "Pomc", "Lepr", "Crh", "Irx3",
                        "Nr5a1", "Kiss1", "Ghrh", "Pirt")
results <- classify_and_assign_cells(seur, genes_of_interest, assay = "Xenium")

table(results$full_classification$Glp1r)
table(results$lepr_positive_assignments$cell_type)

seur <- AddMetaData(seur, results$lepr_positive_assignments %>% column_to_rownames("cell_id"))
```

## 4. Lepr+ neuron subclustering

```{r lepr-subset-cluster}
# Subset to Lepr+ cells
seur_lepr <- subset(seur, cells = colnames(seur)[results$full_classification$Lepr == TRUE])
seur_lepr <- NormalizeData(seur_lepr, assay = "Xenium", margin = 2)
seur_lepr <- ScaleData(seur_lepr)
seur_lepr <- RunPCA(seur_lepr, features = rownames(seur_lepr))
seur_lepr <- FindNeighbors(seur_lepr, dims = 1:30)
seur_lepr <- FindClusters(seur_lepr, resolution = 0.5)
seur_lepr <- RunUMAP(seur_lepr, dims = 1:30)

DimPlot(seur_lepr,
        cells.highlight = WhichCells(seur_lepr, expression = stat3 == "pos"),
        label = TRUE)
```

```{r remove-non-neuronal}
# Remove non-neuronal clusters (identified by marker gene inspection)
seur_lepr_neur <- subset(seur_lepr, idents = c(0, 9, 8, 4, 12), invert = TRUE)

seur_lepr_neur <- NormalizeData(seur_lepr_neur, assay = "Xenium", margin = 2)
seur_lepr_neur <- ScaleData(seur_lepr_neur)
seur_lepr_neur <- RunPCA(seur_lepr_neur, features = rownames(seur_lepr_neur))
seur_lepr_neur <- FindNeighbors(seur_lepr_neur, dims = 1:30)
seur_lepr_neur <- FindClusters(seur_lepr_neur, resolution = 0.5)
seur_lepr_neur <- RunUMAP(seur_lepr_neur, dims = 1:30)

DimPlot(seur_lepr_neur,
        cells.highlight = WhichCells(seur_lepr_neur, expression = stat3 == "pos"),
        label = TRUE)
```

## 5. Cluster marker genes

```{r cluster-markers, fig.width=4, fig.height=4}
marks <- FindAllMarkers(seur_lepr_neur, only.pos = TRUE)
marks %>% group_by(cluster) %>% dplyr::slice(1:2) %>% pull(gene)
```

```{r marker-violin-plots, fig.width=4, fig.height=4}
scCustomize::Stacked_VlnPlot(
  seur_lepr_neur,
  c("Glp1r", "Nr5a1", "Agrp", "Nts", "Pomc", "Irx3", "Tbx19"),
  slot = "counts",
  idents = c(0, 2, 3, 8, 4, 9, 12)
)
```

## 6. UMAP visualization

```{r umap-panels, fig.width=12, fig.height=5}
p <- DimPlot(seur_lepr_neur, label = FALSE, shuffle = TRUE) +
  theme_void() + NoLegend() + coord_equal()
p2 <- LabelClusters(p, id = "ident", fontface = "bold", color = "black", size = 4, repel = FALSE)

p2 + (DimPlot(seur_lepr_neur, group.by = "diet", shuffle = TRUE,
              cells.highlight = WhichCells(seur_lepr_neur, expression = stat3 == "pos"),
              sizes.highlight = 0.2) +
        theme_void() + NoLegend() + coord_equal()) & ggtitle(NULL)

ggsave(file.path(out_dir, "lepr_spatial.pdf"), width = 8, height = 4)
```

## 7. pSTAT3 enrichment analysis

### 7.1 Overall pSTAT3+ proportion by diet

```{r pstat3-overall, fig.width=4, fig.height=3}
model <- seur_lepr_neur[[]] %>%
  group_by(stat3, diet, animal) %>%
  dplyr::count() %>%
  pivot_wider(names_from = stat3, values_from = n, values_fill = 0) %>%
  mutate(tot = neg + pos) %>%
  glm(cbind(pos, neg) ~ diet, data = ., family = binomial)

emmeans::emmeans(model, pairwise ~ diet)

seur_lepr_neur[[]] %>%
  group_by(stat3, diet, section, animal) %>%
  dplyr::count() %>%
  pivot_wider(names_from = stat3, values_from = n) %>%
  mutate(pct = pos / (neg + pos)) %>%
  group_by(animal, diet) %>%
  summarise(mean = mean(pct), .groups = "drop") %>%
  ggplot() +
  aes(x = diet, y = mean) +
  geom_jitter(width = 0.1) +
  stat_summary(geom = "crossbar", aes(group = 1), width = 0.5) +
  geom_signif(comparisons = list(c("dio", "chow")),
              map_signif_level = FALSE, textsize = 0,
              tip_length = c(0.8, 0.01), color = "black") +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "% pSTAT3(+)") +
  ggthemes::theme_clean() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.background = element_blank()) +
  NoLegend()

ggsave(file.path(out_dir, "pstat_alllepr.pdf"), h = 3, w = 2)
```

### 7.2 pSTAT3+ proportion by cluster and diet

```{r pstat3-by-cluster, fig.height=2, fig.width=4}
barplot_data <- seur_lepr_neur[[]] %>%
  filter(seurat_clusters %in% c(0, 2, 3, 4, 8, 9, 12)) %>%
  group_by(seurat_clusters, stat3, animal, diet) %>%
  dplyr::count() %>%
  group_by(animal, seurat_clusters, diet) %>%
  mutate(pct = n / sum(n)) %>%
  filter(stat3 == "pos")

all_combinations <- expand.grid(
  factor = unique(barplot_data[["seurat_clusters"]]),
  group = unique(barplot_data[["diet"]])
)

complete_data <- all_combinations %>%
  full_join(barplot_data, by = c("factor" = "seurat_clusters", "group" = "diet")) %>%
  mutate(value = ifelse(is.na(pct), 1e-5, pct))

complete_data %>%
  ggplot() +
  aes(fct_rev(factor), value, fill = group) +
  stat_summary(geom = "col", position = position_dodge(), color = "black") +
  stat_summary(geom = "errorbar", width = 0.2, position = position_dodge(width = 0.9)) +
  scale_y_continuous(labels = scales::percent) +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  coord_flip()

ggsave(file.path(out_dir, "barplot.pdf"), h = 4, w = 3)
```

### 7.3 Binomial GLM: diet × cluster interaction

```{r pstat3-glm}
seur_lepr_neur[[]] %>%
  filter(seurat_clusters %in% c(0, 2, 3, 4, 8, 9, 12)) %>%
  mutate(seurat_clusters = factor(seurat_clusters)) %>%
  group_by(stat3, diet, section, seurat_clusters, slide, animal) %>%
  dplyr::count() %>%
  ungroup() %>%
  pivot_wider(names_from = stat3, values_from = n) %>%
  mutate(tot = neg + pos) %>%
  lme4::glmer(cbind(pos, tot - pos) ~ diet * seurat_clusters + (1|animal) + (1|slide),
              data = ., family = binomial) %>%
  emmeans::emmeans(., pairwise ~ diet | seurat_clusters) %$%
  contrasts %>%
  data.frame() %>%
  mutate(padj = p.adjust(p.value, method = "BH")) %>%
  arrange(p.value) %>%
  { write.csv(., file.path(out_dir, "pstat3_glm_stats.csv"), row.names = FALSE); . }
```

### 7.4 Alluvial plot

```{r alluvial, fig.width=4, fig.height=3}
seur_lepr_neur[[]] %>%
  group_by(stat3, diet, seurat_clusters) %>%
  dplyr::count() %>%
  ungroup() %>%
  complete(stat3, diet, seurat_clusters, fill = list(n = 1)) %>%
  filter(stat3 == "pos", seurat_clusters %in% c(0, 9, 2, 5, 11, 8, 3)) %>%
  ggplot() +
  aes(x = diet, y = n, fill = seurat_clusters, group = seurat_clusters) +
  geom_col(width = 0.5, color = "white", linewidth = 0.1) +
  geom_flow(aes(alluvium = seurat_clusters), alpha = 0.25, color = "white",
            linewidth = 0.1, curve_type = "linear", width = 0.5) +
  cowplot::theme_minimal_vgrid() +
  theme(panel.grid.major = element_blank()) +
  scale_x_discrete(labels = c("dio" = "HFD", "chow" = "Chow")) +
  labs(x = NULL, y = "# pSTAT3+ Cells")

ggsave(file.path(out_dir, "alluvialbar.pdf"), h = 3, w = 4.5)
```

## 8. Differential expression (diet effect within clusters)

```{r de-analysis}
# Example: DE in cluster 0 (Glp1r+ neurons), diet effect with animal as covariate
marks <- FindMarkers(seur_lepr_neur, ident.1 = "dio", subset.ident = 0,
                     group.by = "diet", slot = "counts",
                     latent.vars = "animal", test.use = "negbinom")

marks %>%
  rownames_to_column("gene") %>%
  ggplot() +
  aes(avg_log2FC, -log10(p_val_adj)) +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = gene)) +
  labs(x = "log2(Fold Change)") +
  ggthemes::theme_clean() +
  theme(plot.background = element_blank()) +
  NoLegend()

ggsave(file.path(out_dir, "glp1r_marks.pdf"), h = 3, w = 5)
```

## 9. Leptin target gene expression by cluster

```{r leptin-targets, fig.height=4, fig.width=4}
scCustomize::Stacked_VlnPlot(
  seur_lepr_neur,
  idents = c(3, 8, 0, 4, 9, 2, 12),
  c("Socs3", "Nlrc5", "Sbno2"),
  split.by = "diet", pt.size = 0, slot = "counts"
)
ggsave(file.path(out_dir, "lep_targs.pdf"), h = 2, w = 5)
```

## Session info

```{r session-info}
sessionInfo()
```
