---
title: "Figure 2 — Lepr/Glp1r neurons progressively adapt to diet-induced obesity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup_libraries_data}
library(Seurat)
library(paletteer)
library(tidyverse)
library(edgeR)
library(ComplexHeatmap)
library(circlize)
library(DESeq2)
library(dynamicTreeCut)
library(clusterProfiler)
library(org.Mm.eg.db)
library(patchwork)
library(magrittr)
library(lme4)
library(emmeans)

source(here::here("R/basic_processing.R"))
source(here::here("R/analysis_wrappers.R"))
source(here::here("R/helper_functions.R"))

BASE_SIZE <- 8

fig_path_2 <- here::here("output_figures/paper/figure2/")
tab_path_2 <- here::here("output_figures/paper/figure2/tables/")

cols_by_geno <- c("mut" = "red", "wt" = "grey70")
cols_by_diet <- c("grey70", "#cb6a49", "#a46cb7", "#7aa457")
diet_cols    <- c("HFD" = "#E8682F", "Chow" = "#F0AD74", "Refeed" = "#58C2AF", "Fast" = "#165C5B")
my_colors    <- paletteer::paletteer_d("lisa::FridaKahlo")

activity_genes <- c("Bdnf", "Crem", "Egr1", "Nr4a1", "Nr4a3",
                    "Pde10a", "Fos", "Fosb", "Fosl2", "Homer1",
                    "Jun", "Junb", "Vgf")

# LGS genes derived from MINT sPLS-DA (see paper/lgs.Rmd)
lgs_genes <- read.csv(here::here("output_figures/paper/lgs/lgs_genes.csv"))
leptin_geneset <- lgs_genes$gene[lgs_genes$value.var < 0]

# Load all required targets
targets::tar_config_set(store = here::here("_targets"))
targets::tar_load(arc_diet_labeled)
targets::tar_load(pb_alldiet)
targets::tar_load(celltypes_diet)
names(pb_alldiet) <- celltypes_diet

targets::tar_load(hfdveh_celldist)
targets::tar_load(fastveh_celldist)

targets::tar_load(arc_lepip)
targets::tar_load(pseudobulk_lepip)
targets::tar_load(celltypes)
names(pseudobulk_lepip) <- celltypes

targets::tar_load(lepip_celldist)
targets::tar_load(obwt_celldist)

# Compute Lepr-expressing clusters — shared across all panels
DefaultAssay(arc_diet_labeled) <- "RNA"
lepr_exp <- AverageExpression(arc_diet_labeled, features = "Lepr", group.by = "seurat_clusters")
lepclus  <- names(as.data.frame(lepr_exp$RNA))[which(lepr_exp$RNA > 1)]
lepclus  <- lepclus[-5]   # drop low-quality cluster (5th element)

# Shared color palette for Lepr clusters across all panels
lepr_pal  <- c("#B85182", "#9A84B7", "#A0C3E7", "#7B9E6B", "#E8B85D", "#D4A27F")[seq_along(lepclus)]
lepr_cols <- setNames(lepr_pal, lepclus)
```

## Figure 2b — UMAP of MBH neurons with Lepr subtypes highlighted

```{r panel_2b_umap}
# Build color vector: Lepr clusters = distinct colors, all others = grey90
all_clusters <- as.character(sort(unique(arc_diet_labeled$seurat_clusters)))
umap_cols    <- setNames(rep("grey90", length(all_clusters)), all_clusters)
umap_cols[names(lepr_cols)] <- lepr_cols

p1 <- DimPlot(arc_diet_labeled, group.by = "seurat_clusters",
              cols = umap_cols, pt.size = 0.1) +
    theme_void() + NoLegend() + ggtitle(NULL)

p2 <- DimPlot(arc_diet_labeled, group.by = "treatment",
              cols = diet_cols, pt.size = 0.1, shuffle = TRUE) +
    theme_void() + NoLegend() + ggtitle(NULL)

p1 + p2
ggsave(paste0(fig_path_2, "figure2b_umap.pdf"), h = 3, w = 6)
```

## Figure 2c — Diet effects on neuronal transcriptomes (scDist)

```{r panel_2c_celldist}
dat_fast <- fastveh_celldist$results %>%
    janitor::clean_names() %>%
    rownames_to_column("clus") %>%
    dplyr::select(clus, dist, x95_percent_ci_low, x95_percent_ci_upper)

dat_hfd <- hfdveh_celldist$results %>%
    janitor::clean_names() %>%
    rownames_to_column("clus") %>%
    dplyr::select(clus, dist, x95_percent_ci_low, x95_percent_ci_upper)

merged_df <- merge(dat_fast, dat_hfd, by = "clus", suffixes = c("_fast", "_hfd")) %>%
    mutate(
        pt_color  = ifelse(clus %in% names(lepr_cols), lepr_cols[clus], "grey70"),
        pt_alpha  = ifelse(clus %in% names(lepr_cols), 1, 0.5),
        clus_label = ifelse(clus %in% names(lepr_cols), clus, "")
    )

ggplot(merged_df, aes(x = dist_fast, y = dist_hfd)) +
    geom_errorbarh(aes(xmin = x95_percent_ci_low_fast, xmax = x95_percent_ci_upper_fast,
                       color = I(pt_color)), height = 0, alpha = 0.3, linewidth = 0.4) +
    geom_errorbar(aes(ymin = x95_percent_ci_low_hfd, ymax = x95_percent_ci_upper_hfd,
                      color = I(pt_color)), width = 0, alpha = 0.3, linewidth = 0.4) +
    geom_point(aes(color = I(pt_color), alpha = I(pt_alpha)), size = 2) +
    geom_abline(lty = 2, linewidth = 0.4) +
    ggrepel::geom_text_repel(aes(label = clus_label),
                             size = 2.5, max.overlaps = 15, show.legend = FALSE) +
    labs(x = "Fasting Effect", y = "HFD Effect") +
    coord_equal() +
    cowplot::theme_cowplot(font_size = BASE_SIZE)

ggsave(paste0(fig_path_2, "figure2c_celldist.pdf"), h = 5, w = 5)
```

## Figure 2d — Leptin gene signature (LGS) across diet conditions

```{r panel_2d_lgs}
DefaultAssay(arc_diet_labeled) <- "RNA"
arc_diet_labeled <- AddModuleScore(arc_diet_labeled, features = list(leptin_geneset), assay = "RNA")

# All Lepr clusters for stats; focal 3 for plotting
all_lepr_clusters <- c("Agrp", "12", "29", "26", "28")
all_lepr_labels   <- c("Agrp" = "AgRP", "12" = "Lepr/Glp1r", "29" = "POMC",
                        "26" = "Tbx19", "28" = "Irx3")
focal_labels <- c("AgRP", "Lepr/Glp1r", "POMC")

lgs_all <- arc_diet_labeled[[]] %>%
    filter(seurat_clusters %in% all_lepr_clusters) %>%
    mutate(ct_label  = all_lepr_labels[as.character(seurat_clusters)],
           ct_label  = factor(ct_label, c("AgRP", "Lepr/Glp1r", "POMC", "Tbx19", "Irx3")),
           treatment = fct_relevel(treatment, c("Fast", "Refeed", "Chow", "HFD"))) %>%
    group_by(hash.mcl.ID, hash_pool, treatment, ct_label) %>%
    summarise(mean_lgs = mean(Cluster1), .groups = "drop")

# Model + contrasts across all Lepr clusters, adjust together
lgs_emm <- lgs_all %>%
    nest(data = -ct_label) %>%
    mutate(
        mod = purrr::map(data, ~lm(mean_lgs ~ treatment, data = .x)),
        emm = purrr::map(mod, ~emmeans::emmeans(.x, pairwise ~ treatment)$contrasts %>% data.frame())
    ) %>%
    unnest(emm) %>%
    ungroup() %>%
    mutate(padj = p.adjust(p.value, "BH")) %>%
    dplyr::select(ct_label, contrast, estimate, SE, p.value, padj)

# Filter to focal 3 for plotting
lgs_focal <- lgs_all %>% filter(ct_label %in% focal_labels)

write_csv(lgs_emm, paste0(tab_path_2, "lgs_focal_diet_stats.csv"))

# Consecutive comparisons only: Fast-Refeed, Refeed-Chow, Chow-HFD
consecutive_contrasts <- c("Fast - Refeed", "Refeed - Fast",
                           "Refeed - Chow", "Chow - Refeed",
                           "Chow - HFD", "HFD - Chow")

sig_label <- function(p) case_when(p < 0.001 ~ "***", p < 0.01 ~ "**", p < 0.05 ~ "*", p < 0.1 ~ "+", TRUE ~ "ns")

# Build ggsignif annotation data from emmeans results
signif_df <- lgs_emm %>%
    filter(ct_label %in% focal_labels, contrast %in% consecutive_contrasts) %>%
    mutate(
        xmin = case_when(grepl("Fast.*Refeed|Refeed.*Fast", contrast) ~ "Fast",
                         grepl("Refeed.*Chow|Chow.*Refeed", contrast) ~ "Refeed",
                         grepl("Chow.*HFD|HFD.*Chow", contrast) ~ "Chow"),
        xmax = case_when(grepl("Fast.*Refeed|Refeed.*Fast", contrast) ~ "Refeed",
                         grepl("Refeed.*Chow|Chow.*Refeed", contrast) ~ "Chow",
                         grepl("Chow.*HFD|HFD.*Chow", contrast) ~ "HFD"),
        label = sig_label(padj)
    ) %>%
    filter(label != "ns") %>%
    distinct(ct_label, xmin, xmax, label, padj)

# Compute y positions per facet
y_max_df <- lgs_focal %>%
    group_by(ct_label) %>%
    summarise(y_max = max(mean_lgs), .groups = "drop")
y_range <- diff(range(lgs_focal$mean_lgs))

signif_df <- signif_df %>%
    left_join(y_max_df, by = "ct_label") %>%
    group_by(ct_label) %>%
    mutate(y_pos = y_max + y_range * 0.05 * row_number()) %>%
    ungroup()

p_lgs <- ggplot(lgs_focal, aes(treatment, mean_lgs, fill = treatment)) +
    geom_boxplot(outlier.shape = NA, width = 0.6, linewidth = 0.35) +
    geom_jitter(width = 0.15, size = 0.5, alpha = 0.5, shape = 16) +
    facet_wrap(~ct_label, nrow = 1, scales = "free_y") +
    scale_fill_manual(values = diet_cols) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(axis.text.x    = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.background = element_blank()) +
    labs(x = NULL, y = "Leptin Gene Signature")

if (nrow(signif_df) > 0) {
    p_lgs <- p_lgs +
        ggsignif::geom_signif(
            data = signif_df, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, y_position = y_pos, annotations = label),
            manual = TRUE, textsize = 2.5, tip_length = 0.01, vjust = 0.5
        )
}

ggsave(paste0(fig_path_2, "figure2d_lgs.pdf"), plot = p_lgs, h = 3, w = 5)
```

## Figure 2d (all clusters) — LGS across all Lepr clusters

```{r panel_2d_lgs_all_clusters}
# Consecutive contrasts for all clusters
consecutive_all <- c("Fast - Refeed", "Refeed - Fast",
                     "Refeed - Chow", "Chow - Refeed",
                     "Chow - HFD", "HFD - Chow")

signif_all <- lgs_emm %>%
    filter(contrast %in% consecutive_all) %>%
    mutate(
        xmin = case_when(grepl("Fast.*Refeed|Refeed.*Fast", contrast) ~ "Fast",
                         grepl("Refeed.*Chow|Chow.*Refeed", contrast) ~ "Refeed",
                         grepl("Chow.*HFD|HFD.*Chow", contrast) ~ "Chow"),
        xmax = case_when(grepl("Fast.*Refeed|Refeed.*Fast", contrast) ~ "Refeed",
                         grepl("Refeed.*Chow|Chow.*Refeed", contrast) ~ "Chow",
                         grepl("Chow.*HFD|HFD.*Chow", contrast) ~ "HFD"),
        label = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", padj < 0.1 ~ "+", TRUE ~ "ns")
    ) %>%
    filter(label != "ns") %>%
    distinct(ct_label, xmin, xmax, label, padj)

# Y positions per facet
y_max_all <- lgs_all %>%
    group_by(ct_label) %>%
    summarise(y_max = max(mean_lgs), .groups = "drop")
y_range_all <- diff(range(lgs_all$mean_lgs))

signif_all <- signif_all %>%
    left_join(y_max_all, by = "ct_label") %>%
    group_by(ct_label) %>%
    mutate(y_pos = y_max + y_range_all * 0.05 * row_number()) %>%
    ungroup()

p_lgs_all <- ggplot(lgs_all, aes(treatment, mean_lgs, fill = treatment)) +
    geom_boxplot(outlier.shape = NA, width = 0.6, linewidth = 0.35) +
    geom_jitter(width = 0.15, size = 0.5, alpha = 0.5, shape = 16) +
    facet_wrap(~ct_label, nrow = 1, scales = "free_y") +
    scale_fill_manual(values = diet_cols) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.background = element_blank()) +
    labs(x = NULL, y = "Leptin Gene Signature")

if (nrow(signif_all) > 0) {
    p_lgs_all <- p_lgs_all +
        ggsignif::geom_signif(
            data = signif_all, inherit.aes = FALSE,
            aes(xmin = xmin, xmax = xmax, y_position = y_pos, annotations = label),
            manual = TRUE, textsize = 2.5, tip_length = 0.01, vjust = 0.5
        )
}

ggsave(paste0(fig_path_2, "figure2d_lgs_all_clusters.pdf"), plot = p_lgs_all, h = 3, w = 8)

# Print full stats table for reference
cat("\n=== LGS stats — all Lepr clusters, all pairwise contrasts ===\n")
lgs_emm %>%
    arrange(ct_label, contrast) %>%
    mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**", padj < 0.05 ~ "*", TRUE ~ "")) %>%
    print(n = 50)

# Key contrasts for manuscript: Chow vs HFD and Chow vs Fast per cluster
cat("\n=== Key contrasts: Chow vs HFD ===\n")
lgs_emm %>%
    filter(grepl("Chow.*HFD|HFD.*Chow|Chow.*Fast|Fast.*Chow", contrast)) %>%
    arrange(ct_label, contrast) %>%
    print(n = 20)
```

## Figure 2e — Hierarchical clustering of DEGs in Lepr/Glp1r neurons

```{r panel_2e_deg_analysis}
pb_objects  <- list()
fit_objects <- list()

for (ct in lepclus) {
    pb_objects[[ct]] <- edger_prep(
        subset(arc_diet_labeled, subset = hash.mcl.ID != "REFED2"),
        celltype_column = "seurat_clusters", trt_group = "treatment", celltypes = ct
    )
}

deg_results <- list()

for (ct in lepclus) {
    keep <- filterByExpr(pb_objects[[ct]], group = pb_objects[[ct]]$samples$treatment,
                         min.count = 5, min.total.count = 15)
    pb_objects[[ct]] <- pb_objects[[ct]][keep, , keep.lib.sizes = FALSE]
    pb_objects[[ct]] <- normLibSizes(pb_objects[[ct]])
    pb_objects[[ct]]$samples$treatment <- factor(pb_objects[[ct]]$samples$treatment,
                                                 levels = c("Chow", "HFD", "Refeed", "Fast"))
    design <- model.matrix(~treatment + hash_pool, data = pb_objects[[ct]]$samples)
    colnames(design) <- gsub(":", "_", colnames(design))
    pb_objects[[ct]] <- estimateDisp(pb_objects[[ct]], design, robust = TRUE)
    fit <- glmQLFit(pb_objects[[ct]], design, robust = TRUE)
    fit_objects[[ct]] <- fit   # save fit for per-contrast extraction
    qlf <- glmQLFTest(fit, coef = 2:4)
    deg_results[[ct]] <- topTags(qlf, n = Inf, p = 0.05)
}

alldeg <- purrr::map_dfr(deg_results, data.frame, .id = "ct")
write_csv(alldeg, paste0(tab_path_2, "diet_degenes.csv"))
cat("Total diet DEGs across Lepr clusters (FDR < 0.05):", nrow(alldeg), "\n")
cat("DEGs in Lepr/Glp1r cluster 12:", nrow(alldeg[alldeg$ct == "12", ]), "\n")
```

## Pearson's r — Fast vs DIO logFC anti-correlation in Lepr/Glp1r neurons (cluster 12)

```{r panel_2e_pearson, fig.width=8, fig.height=3}
# Scatter: HFD logFC (x) vs Fasting logFC (y) for AgRP, Lepr/Glp1r, POMC
# Uses F-test DEGs per cluster (genes responsive to any nutritional state)
cor_clusters <- c("Agrp", "12", "29")
cor_labels <- c("Agrp" = "AgRP", "12" = "Lepr/Glp1r", "29" = "POMC")

cor_plot_data <- list()
cor_stats <- list()

for (ct in cor_clusters) {
    ct_degs <- alldeg %>% filter(ct == !!ct) %>% pull(gene)
    if (length(ct_degs) < 10) next

    fast_fc <- alldeg %>% filter(ct == !!ct) %>%
        dplyr::select(gene, logFC_fast = logFC.treatmentFast)
    hfd_fc <- alldeg %>% filter(ct == !!ct) %>%
        dplyr::select(gene, logFC_hfd = logFC.treatmentHFD)

    merged <- inner_join(fast_fc, hfd_fc, by = "gene") %>%
        mutate(ct_label = cor_labels[ct])
    cor_plot_data[[ct]] <- merged

    ct_cor <- cor.test(merged$logFC_hfd, merged$logFC_fast, method = "pearson")
    cor_stats[[ct]] <- data.frame(
        ct_label = cor_labels[ct],
        r = round(ct_cor$estimate, 3),
        p = signif(ct_cor$p.value, 3),
        n = nrow(merged)
    )
}

cor_all <- bind_rows(cor_plot_data) %>%
    mutate(ct_label = factor(ct_label, levels = c("AgRP", "Lepr/Glp1r", "POMC")))
cor_stats_df <- bind_rows(cor_stats) %>%
    mutate(ct_label = factor(ct_label, levels = c("AgRP", "Lepr/Glp1r", "POMC")),
           label = sprintf("r = %.3f\np = %.2e", r, p))

p_fvc_hvc <- ggplot(cor_all, aes(logFC_hfd, logFC_fast)) +
    geom_point(size = 0.3, alpha = 0.3, color = "grey40") +
    geom_smooth(method = "lm", color = "#C0392B", se = FALSE, linewidth = 0.5) +
    geom_hline(yintercept = 0, lty = 2, color = "grey60", linewidth = 0.3) +
    geom_vline(xintercept = 0, lty = 2, color = "grey60", linewidth = 0.3) +
    geom_text(data = cor_stats_df,
              aes(x = -Inf, y = Inf, label = label),
              hjust = -0.1, vjust = 1.2, size = 2.5, color = "#C0392B") +
    facet_wrap(~ct_label, nrow = 1, scales = "free") +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(strip.background = element_blank()) +
    labs(x = "log2FC (HFD vs Chow)", y = "log2FC (Fasting vs Chow)")

ggsave(paste0(fig_path_2, "figure2e_fast_hfd_correlation.pdf"), plot = p_fvc_hvc, h = 3, w = 8)

# Print stats
cat("Pearson correlations (Fasting vs HFD logFC, diet-regulated DEGs):\n")
print(cor_stats_df)
```

```{r panel_2e_glp1r_heatmap}
pbd <- pb_objects[["12"]]
pbd$samples$treatment <- factor(pbd$samples$treatment, levels = c("Fast", "Refeed", "Chow", "HFD"))

logcpm <- edgeR::cpm(pbd, prior.count = 5, log = T)
logcpm <- limma::removeBatchEffect(logcpm, pbd$samples$hash_pool,
                                   design = model.matrix(~treatment, data = pbd$samples))

glp_deg  <- alldeg %>% filter(ct %in% 12) %>% pull(gene)
expr_dat <- logcpm[rownames(logcpm) %in% glp_deg, ]
expr_dat <- t(scale(t(expr_dat)))

# Genes to label — neuronal activity, GABA receptors, and leptin-response markers
tolabel <- c("Nr4a3", "Npas4", "Gabra3", "Gabra4", "Gabra5", "Grip1", "Gabra1", "Sv2c", "Pde10a", "Grin1",
             "Vgf", "Tenm2", "Itpr1", "Nrcam", "Scn9a", "Fos", "Gabrb1", "Gabrb2", "Gabrb3",
             "Nr4a1", "Junb", "Stat3", "Bcl3", "Cxcl12")
idx       <- match(tolabel, rownames(expr_dat))
valid_idx <- idx[!is.na(idx)]
valid_lab <- tolabel[!is.na(idx)]

ha <- HeatmapAnnotation(
    d = pbd$samples$treatment[order(pbd$samples$treatment)],
    col = list(d = c("HFD" = my_colors[[1]], "Chow" = my_colors[[2]],
                     "Refeed" = my_colors[[3]], "Fast" = my_colors[[4]])),
    simple_anno_size = unit(0.3, "cm"), annotation_name_gp = gpar(fontsize = BASE_SIZE))
ra <- rowAnnotation(foo = anno_mark(at = valid_idx, labels = valid_lab,
                                    labels_gp = gpar(fontsize = BASE_SIZE)))

pdf(paste0(fig_path_2, "figure2e_glp1r_hmap.pdf"), h = 4, w = 5)
set.seed(426)
hmap <- Heatmap(expr_dat[, order(pbd$samples$treatment)],
    show_row_dend = T, show_row_names = F,
    right_annotation = ra, row_split = 5,
    top_annotation = ha, cluster_columns = F,
    clustering_method_rows = "ward.D2",
    show_column_names = F, row_names_gp = gpar(fontsize = 2))
ht <- draw(hmap)
dev.off()

cluster_genes <- purrr::map(row_order(ht), function(x) rownames(expr_dat)[x])
names(cluster_genes) <- names(row_order(ht))
cluster_genes <- enframe(cluster_genes) %>% unnest(value)
write_csv(cluster_genes, paste0(tab_path_2, "glp1r_lepr_hmap_genes.csv"))
cat("Total genes in diet heatmap (cluster 12):", nrow(cluster_genes), "\n")
```

```{r panel_2e_go_enrichment}
compare_result <- compareCluster(value ~ name,
    data = cluster_genes, maxGSSize = 300, minGSSize = 5,
    fun = "enrichGO",
    OrgDb = org.Mm.eg.db,
    ont = "ALL",
    keyType = "SYMBOL",
    pvalueCutoff = 0.05,
    universe = rownames(pbd$counts))

compare_result@compareClusterResult %>% write_csv(paste0(tab_path_2, "glp1r_lepr_hmap_goterms.csv"))

topgo <- simplify(compare_result, cutoff = 0.5, by = "p.adjust",
                  select_fun = min, measure = "Wang", semData = NULL)

dotplot(topgo)
ggsave(paste0(fig_path_2, "figure2e_goterms.pdf"), h = 10, w = 6)
```

## Figure 2f — Experimental design schematic (assembled in Illustrator)

<!-- 2f is a schematic of the leptin IP experiment: vehicle n=22, leptin n=20, 1-24h
     time points. No code required — assembled externally. -->

## Figure 2g — Leptin response per cell type (scDist)

```{r panel_2g_lepip_celldist}
lepip_cd_data <- lepip_celldist$results %>%
    rownames_to_column("row") %>%
    separate(row, into = c("ct", "comp")) %>%
    janitor::clean_names() %>%
    dplyr::select(comp, ct, dist, x95_percent_ci_low, x95_percent_ci_upper, p_val) %>%
    pivot_wider(
        names_from  = comp,
        values_from = c(dist, x95_percent_ci_low, x95_percent_ci_upper, p_val)
    ) %>%
    filter(ct %in% names(which(table(arc_lepip$predicted.celltype) > 500))) %>%
    mutate(
        pt_color = case_when(
            ct %in% names(lepr_cols) ~ lepr_cols[ct],
            TRUE ~ "grey70"
        )
    )

lepip_cd_data %>%
    ggplot(aes(fct_reorder(ct, -dist_wt), dist_wt)) +
    geom_col(aes(fill = I(pt_color),
                 color = I(ifelse(p_val_wt < 0.05, "black", "white"))),
             linewidth = 0.3) +
    geom_errorbar(aes(ymin = x95_percent_ci_low_wt, ymax = x95_percent_ci_upper_wt),
                  width = 0.3, linewidth = 0.4) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    NoLegend() +
    labs(x = NULL, y = "Transcriptome Distance (Leptin vs Saline)")

ggsave(paste0(fig_path_2, "figure2g_lepip_celldist.pdf"), h = 3, w = 8)
```

## Figure 2g2 — LGS response to leptin IP over time

```{r panel_2g2_lgs_lepip_timecourse}
# Score LGS on leptin IP dataset
arc_lepip <- AddModuleScore(arc_lepip, features = list(leptin_geneset), assay = "RNA")

# Summarise per animal, cell type, timepoint — all Lepr clusters
lgs_lepip_all <- arc_lepip[[]] %>%
    filter(dataset == "lepip" & predicted.celltype %in% c("12", "Agrp", "29", "26", "28"),
           geno == "wt") %>%
    mutate(treatment = factor(treatment, c("Lep", "Sal")),
           ct_label = all_lepr_labels[as.character(predicted.celltype)],
           ct_label = factor(ct_label, c("AgRP", "Lepr/Glp1r", "POMC", "Tbx19", "Irx3"))) %>%
    group_by(treatment, ct_label, hash.mcl.ID, time) %>%
    summarise(mean_lgs = mean(Cluster1), .groups = "drop")

# Linear model per cell type: LGS ~ treatment * time, adjust across all clusters
lgs_lepip_emm <- lgs_lepip_all %>%
    nest(data = -ct_label) %>%
    mutate(
        mod = purrr::map(data, ~lm(mean_lgs ~ treatment * factor(time), data = .x)),
        emm = purrr::map(mod, ~emmeans::emmeans(.x, pairwise ~ treatment | time)$contrasts %>% data.frame())
    ) %>%
    unnest(emm) %>%
    ungroup() %>%
    mutate(padj = p.adjust(p.value, "BH"))

write_csv(lgs_lepip_emm, paste0(tab_path_2, "lgs_lepip_timecourse_stats.csv"))

# Plot: effect size (Lep - Sal) over time — focal 3 only
lgs_lepip_plot <- lgs_lepip_emm %>%
    filter(ct_label %in% focal_labels) %>%
    mutate(sig = padj < 0.05)

p_lgs_time <- ggplot(lgs_lepip_plot) +
    aes(x = factor(time, levels = c("1", "3", "6", "24")), y = estimate,
        color = ct_label, group = ct_label) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    geom_line() +
    geom_errorbar(aes(ymin = estimate - SE, ymax = estimate + SE), width = 0) +
    geom_point(data = ~filter(.x, !sig), shape = 21, size = 2, fill = "white", stroke = 0.5) +
    geom_point(data = ~filter(.x, sig), aes(fill = ct_label), shape = 21, size = 2, stroke = 0.5) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    labs(x = "Hours post-injection", y = "LGS (Leptin - Saline)") +
    NoLegend()

ggsave(paste0(fig_path_2, "figure2g2_lgs_lepip_timecourse.pdf"), plot = p_lgs_time, h = 3, w = 4)
```

## Figure 2h — Leptin-regulated gene heatmaps (AgRP and Lepr/Glp1r neurons)

```{r panel_2h_agrp_timecourse}
pb <- pseudobulk_lepip[["Agrp"]]
pb <- pb[, pb$samples$dataset == "lepip" & pb$samples$time != 24 & pb$samples$geno == "wt"]
colnames(pb) <- paste0(colnames(pb), "_", pb$samples$time)
pb$samples$treatment <- factor(pb$samples$treatment, levels = c("Sal", "Lep"))
pb$samples$time      <- factor(pb$samples$time, levels = c("1", "3", "6"))

design <- model.matrix(~treatment + time + hash_pool, data = pb$samples)
colnames(design) <- gsub(":", "_", gsub("[()]", "", colnames(design)))

fit <- voomLmFit(pb, design)
fit2 <- eBayes(fit, robust = TRUE)

agrp_lep_all <- topTable(fit2, coef = "treatmentLep", n = Inf)
write_csv(agrp_lep_all %>% { if (!"gene" %in% names(.)) rownames_to_column(., "gene") else . },
          paste0(tab_path_2, "agrp_leptin_degs_wt.csv"))
cat("AgRP WT leptin DEGs (FDR < 0.05):", sum(agrp_lep_all$adj.P.Val < 0.05), "\n")

alldeg_agrp <- topTable(fit2, coef = "treatmentLep", n = Inf, p = 0.1)
filtdeg_agrp <- alldeg_agrp[abs(alldeg_agrp$logFC) > 0.5, ]

logcpm_agrp <- edgeR::cpm(pb, prior.count = 5, log = T)
logcpm_agrp <- limma::removeBatchEffect(logcpm_agrp, batch = pb$samples$hash_pool,
    design = model.matrix(~treatment + time, data = pb$samples))

expr_data <- t(scale(t(logcpm_agrp[rownames(logcpm_agrp) %in% rownames(filtdeg_agrp), ])))
gene_hclust <- hclust(dist(expr_data), method = "ward.D2")
clusters <- cutree(gene_hclust, k = 4)

df <- bind_cols(data.frame(expr_data, check.names = F), enframe(clusters)) %>%
    pivot_longer(c(-name, -value), names_to = "sample", values_to = "expression") %>%
    separate(sample, into = c("hash.mcl.ID", "treatment", "time"), sep = "_")

agrp_timeplot <- df %>%
    group_by(value, time, treatment, hash.mcl.ID) %>%
    summarise(mean = mean(expression)) %>%
    ggplot() +
    aes(factor(time, c(1, 3, 6, 24)), mean,
        group = treatment, color = treatment) +
    stat_summary(geom = "point", size = 0.5) +
    stat_summary(geom = "line", linewidth = 0.25) +
    facet_wrap(~value, ncol = 1, scales = "free_y", switch = "y") +
    scale_color_manual(values = c("Lep" = "red", "Sal" = "black")) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(strip.text       = element_text(colour = "black"),
          strip.background = element_rect(fill = "white"),
          strip.placement  = "outside") +
    labs(x = NULL, y = NULL, color = NULL) +
    scale_x_discrete(expand = c(0.05, 0))

ggsave(paste0(fig_path_2, "figure2h_agrp_timecourse.pdf"), plot = agrp_timeplot, h = 5, w = 2)
```

```{r panel_2h_glp1r_lepip_heatmap}
pb_glp <- pseudobulk_lepip[["12"]]
pb_glp <- pb_glp[, pb_glp$samples$dataset == "lepip" & pb_glp$samples$time != 24 & pb_glp$samples$geno == "wt"]
colnames(pb_glp) <- paste0(colnames(pb_glp), "_", pb_glp$samples$time)
pb_glp$samples$treatment <- factor(pb_glp$samples$treatment, levels = c("Sal", "Lep"))
pb_glp$samples$time      <- factor(pb_glp$samples$time, levels = c("1", "3", "6"))

design_glp <- model.matrix(~treatment + time + hash_pool, data = pb_glp$samples)
colnames(design_glp) <- gsub(":", "_", gsub("[()]", "", colnames(design_glp)))

fit_glp  <- voomLmFit(pb_glp, design_glp)
fit2_glp <- eBayes(fit_glp, robust = TRUE)

glp1r_lep_all <- topTable(fit2_glp, coef = "treatmentLep", n = Inf)
write_csv(glp1r_lep_all %>% { if (!"gene" %in% names(.)) rownames_to_column(., "gene") else . },
          paste0(tab_path_2, "glp1r_leptin_degs_wt.csv"))
cat("Lepr/Glp1r WT leptin DEGs (FDR < 0.05):", sum(glp1r_lep_all$adj.P.Val < 0.05), "\n")

alldeg_glp  <- topTable(fit2_glp, coef = "treatmentLep", n = Inf, p = 0.1)
filtdeg_glp <- alldeg_glp[abs(alldeg_glp$logFC) > 0.5, ]

logcpm_glp <- edgeR::cpm(pb_glp, prior.count = 5, log = T)
logcpm_glp <- limma::removeBatchEffect(logcpm_glp, batch = pb_glp$samples$hash_pool,
    design = model.matrix(~treatment + time, data = pb_glp$samples))

expr_glp <- t(scale(t(logcpm_glp[rownames(logcpm_glp) %in% rownames(filtdeg_glp), ])))

tolabel_glp <- c("Fos", "Fosb", "Nr4a1", "Nr4a3", "Socs3", "Stat3", "Egr1", "Vgf", "Npy", "Pomc")
idx_glp <- match(tolabel_glp, rownames(expr_glp))
valid   <- !is.na(idx_glp)
ra_glp  <- rowAnnotation(foo = anno_mark(at = idx_glp[valid], labels = tolabel_glp[valid],
                                         labels_gp = gpar(fontsize = BASE_SIZE)))

pdf(paste0(fig_path_2, "figure2h_glp1r_hmap.pdf"), h = 4, w = 5)
set.seed(426)
Heatmap(expr_glp[, order(pb_glp$samples$treatment)],
    show_row_dend = T, show_row_names = F,
    right_annotation = ra_glp, row_split = 4,
    cluster_columns = F, clustering_method_rows = "ward.D2",
    show_column_names = F, row_names_gp = gpar(fontsize = 2))
dev.off()
```

---

## Supplemental — Leptin IP DE gene counts per cluster (WT only)

```{r lepip_de_all_clusters}
# Run DE for all clusters (WT only, treatment + time + hash_pool where possible)
lepip_de_counts <- list()

for (ct in celltypes) {
    pb <- pseudobulk_lepip[[ct]]
    keep <- pb$samples$dataset == "lepip" & pb$samples$time != 24 & pb$samples$geno == "wt"
    if (sum(keep) < 6) next
    pb_sub <- pb[, keep]
    colnames(pb_sub) <- make.unique(colnames(pb_sub))
    pb_sub$samples$treatment <- factor(pb_sub$samples$treatment, levels = c("Sal", "Lep"))
    pb_sub$samples$time <- factor(pb_sub$samples$time)

    # Try with hash_pool; fall back without if rank-deficient
    design <- tryCatch({
        d <- model.matrix(~ treatment + time + hash_pool, data = pb_sub$samples)
        if (qr(d)$rank < ncol(d)) stop("rank deficient")
        d
    }, error = function(e) model.matrix(~ treatment + time, data = pb_sub$samples))

    tt <- tryCatch({
        f <- voomLmFit(pb_sub, design)
        f <- eBayes(f, robust = TRUE)
        topTable(f, coef = "treatmentLep", n = Inf)
    }, error = function(e) NULL)
    if (is.null(tt)) next

    lepip_de_counts[[ct]] <- data.frame(
        cell_type = ct,
        n_up = sum(tt$adj.P.Val < 0.05 & tt$logFC > 0, na.rm = TRUE),
        n_down = sum(tt$adj.P.Val < 0.05 & tt$logFC < 0, na.rm = TRUE)
    )
}

lepip_de_df <- bind_rows(lepip_de_counts) %>%
    mutate(total = n_up + n_down) %>%
    arrange(desc(total))

write_csv(lepip_de_df, paste0(tab_path_2, "lepip_de_counts_all_clusters.csv"))
cat("Leptin IP DE gene counts (WT only, all clusters):\n")
print(lepip_de_df %>% filter(total > 0))
```

## Supplemental — Leptin IP UMAP overlay

```{r panel_supp_lepip_umap}
# Use ref.umap if available (reference-based projection), otherwise fall back to umap
diet_umap_key  <- intersect(c("ref.umap", "umap"), names(arc_diet_labeled@reductions))[1]
lepip_umap_key <- intersect(c("ref.umap", "umap"), names(arc_lepip@reductions))[1]

diet_coords <- data.frame(arc_diet_labeled@reductions[[diet_umap_key]]@cell.embeddings)
colnames(diet_coords)[1:2] <- c("U1", "U2")
diet_coords$diet    <- arc_diet_labeled$treatment
diet_coords$dataset <- "diet"

lepip_coords <- data.frame(arc_lepip@reductions[[lepip_umap_key]]@cell.embeddings) %>%
    dplyr::sample_frac(0.2)
colnames(lepip_coords)[1:2] <- c("U1", "U2")
lepip_coords$diet    <- NA
lepip_coords$dataset <- "lepip"

mapping_df <- bind_rows(diet_coords, lepip_coords)

bg_plot <- ggplot(mapping_df[mapping_df$dataset == "lepip", ]) +
    aes(U1, U2) +
    geom_point(shape = ".", alpha = 0.1, color = "grey70") +
    theme_void() + coord_equal()

bg_plot +
    geom_point(data = mapping_df[!is.na(mapping_df$diet) & mapping_df$diet == "HFD", ],
               color = cols_by_diet[2], size = 0.15) +
    geom_point(data = mapping_df[!is.na(mapping_df$diet) & mapping_df$diet == "Fast", ],
               color = cols_by_diet[1], size = 0.15) +
    geom_point(data = mapping_df[!is.na(mapping_df$diet) & mapping_df$diet == "Refeed", ],
               color = cols_by_diet[3], size = 0.15)

ggsave(paste0(fig_path_2, "supp_lepip_umap.pdf"), h = 5, w = 5)
```

## Figure 2i — HFD vs Leptin gene overlap in Lepr/Glp1r neurons

```{r panel_supp_hfd_lep_overlap}
# Self-contained: rebuilds both models from scratch so this chunk runs independently

# --- Diet: HFD vs Chow in cluster 12 ---
pb_ov_diet <- edger_prep(
    subset(arc_diet_labeled, subset = hash.mcl.ID != "REFED2"),
    celltype_column = "seurat_clusters", trt_group = "treatment", celltypes = "12"
)
pb_ov_diet$samples$treatment <- factor(pb_ov_diet$samples$treatment,
                                       levels = c("Chow", "HFD", "Refeed", "Fast"))
design_ov  <- model.matrix(~treatment + hash_pool, data = pb_ov_diet$samples)
pb_ov_diet <- estimateDisp(pb_ov_diet, design_ov, robust = TRUE)
fit_ov     <- glmQLFit(pb_ov_diet, design_ov, robust = TRUE)
qlf_hfd    <- glmQLFTest(fit_ov, coef = 2)   # coef 2 = treatmentHFD vs Chow
diet_all   <- topTags(qlf_hfd, n = Inf)$table %>%
    dplyr::select(gene, logFC_hfd = logFC, FDR_hfd = FDR)

# --- Leptin: WT-only response in cluster 12 ---
pb_ov_lep <- pseudobulk_lepip[["12"]]
pb_ov_lep <- pb_ov_lep[, pb_ov_lep$samples$dataset == "lepip" & pb_ov_lep$samples$time != 24 & pb_ov_lep$samples$geno == "wt"]
colnames(pb_ov_lep) <- paste0(colnames(pb_ov_lep), "_", pb_ov_lep$samples$time)
pb_ov_lep$samples$treatment <- factor(pb_ov_lep$samples$treatment, levels = c("Sal", "Lep"))
pb_ov_lep$samples$time      <- factor(pb_ov_lep$samples$time, levels = c("1", "3", "6"))
design_lep <- model.matrix(~treatment + time + hash_pool, data = pb_ov_lep$samples)
colnames(design_lep) <- gsub(":", "_", gsub("[()]", "", colnames(design_lep)))
fit_lep    <- voomLmFit(pb_ov_lep, design_lep)
fit2_lep   <- eBayes(fit_lep, robust = TRUE)
lep_all    <- topTable(fit2_lep, coef = "treatmentLep", n = Inf) %>%
    dplyr::select(gene, logFC_lep = logFC, FDR_lep = adj.P.Val)

# --- Merge with directional consistency ---
overlap_df <- inner_join(diet_all, lep_all, by = "gene") %>%
    mutate(
        same_dir  = sign(logFC_hfd) == sign(logFC_lep),
        sig_group = case_when(
            FDR_hfd < 0.05 & FDR_lep < 0.05 & same_dir  ~ "Concordant",
            FDR_hfd < 0.05 & FDR_lep < 0.05 & !same_dir ~ "Discordant",
            FDR_hfd < 0.05                               ~ "Diet only",
            FDR_lep < 0.05                               ~ "Leptin only",
            TRUE                                         ~ "Neither"
        )
    )

write_csv(overlap_df, paste0(tab_path_2, "hfd_lep_overlap_clus12.csv"))

# --- Pearson correlation across all expressed genes ---
cor_test  <- cor.test(overlap_df$logFC_hfd, overlap_df$logFC_lep, method = "pearson")
cor_label <- sprintf("r = %.2f, p = %.2e", cor_test$estimate, cor_test$p.value)

# Genes to label: activity genes (IEGs) and canonical leptin targets
label_activity <- c("Fos", "Fosb", "Jun", "Junb", "Vgf", "Homer1", "Nr4a1", "Egr1")
label_leptin   <- c("Socs3", "Atf3", "Etv6", "Sbno2", "Nlrc5", "Stat3", "Cish", "Bcl3")
label_gaba     <- c("Gabra3", "Gabra4", "Gabra5", "Gabrb1", "Gabrb2", "Gabrb3")

overlap_df <- overlap_df %>%
    mutate(
        label_class = case_when(
            gene %in% label_activity ~ "Activity",
            gene %in% label_leptin   ~ "Leptin target",
            gene %in% label_gaba     ~ "GABA receptor",
            TRUE ~ NA_character_
        ),
        label = ifelse(!is.na(label_class) & sig_group == "Concordant", gene, NA_character_)
    )

# Label colors by gene class
label_cols <- c("Activity" = "#E74C3C", "Leptin target" = "#2980B9", "GABA receptor" = "#27AE60")

# Restrict to HFD DE genes only
hfd_sig_df <- overlap_df %>% filter(FDR_hfd < 0.05)

# Recalculate Pearson on HFD DE genes only
cor_test_hfd <- cor.test(hfd_sig_df$logFC_hfd, hfd_sig_df$logFC_lep, method = "pearson")
cor_label_hfd <- sprintf("r = %.2f, p = %.2e, n = %d", cor_test_hfd$estimate, cor_test_hfd$p.value, nrow(hfd_sig_df))

highlight_df <- hfd_sig_df %>% filter(!is.na(label_class))

p_overlap <- ggplot(hfd_sig_df, aes(logFC_hfd, logFC_lep)) +
    geom_hline(yintercept = 0, lty = 2, color = "grey50", linewidth = 0.3) +
    geom_vline(xintercept = 0, lty = 2, color = "grey50", linewidth = 0.3) +
    geom_point(data = ~filter(.x, sig_group != "Concordant"),
               color = "grey60", size = 0.6, alpha = 0.5) +
    geom_point(data = ~filter(.x, sig_group == "Concordant"),
               color = "#C0392B", size = 0.8, alpha = 0.6) +
    geom_smooth(method = "lm", color = "black", se = FALSE, linewidth = 0.5) +
    geom_point(data = highlight_df,
               aes(fill = label_class), shape = 21, size = 2.5, stroke = 0.4, color = "black") +
    scale_fill_manual(values = label_cols) +
    ggrepel::geom_text_repel(
        data = highlight_df,
        aes(label = gene, color = label_class),
        size = 2, fontface = "italic", max.overlaps = 25,
        min.segment.length = 0, segment.size = 0.3,
        show.legend = FALSE
    ) +
    scale_color_manual(values = label_cols) +
    annotate("text", x = -Inf, y = Inf, label = cor_label_hfd,
             hjust = -0.1, vjust = 1.3, size = 2.5) +
    annotate("text", x = Inf, y = -Inf,
             label = paste0(sum(hfd_sig_df$sig_group == "Concordant"), " concordant / ",
                            nrow(hfd_sig_df), " HFD DE genes"),
             hjust = 1.1, vjust = -0.5, size = 2, color = "grey40") +
    labs(x = "log2FC (HFD vs Chow)", y = "log2FC (Leptin vs Saline)",
         fill = NULL) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(legend.position = "bottom", legend.text = element_text(size = BASE_SIZE - 1))

ggsave(paste0(fig_path_2, "figure2i_hfd_lep_overlap.pdf"), plot = p_overlap, h = 4, w = 4.5)

# --- Fisher's exact test: concordant overlap only ---
cont_table <- with(overlap_df, table(
    HFD_sig        = FDR_hfd < 0.05,
    Leptin_concord = FDR_lep < 0.05 & same_dir
))
fish_res <- fisher.test(cont_table)

cat("Fisher's exact test — concordant HFD x Leptin DEGs in Lepr/Glp1r neurons\n")
print(cont_table)
cat(sprintf("Odds ratio: %.2f  (95%% CI: %.2f-%.2f)\n",
            fish_res$estimate, fish_res$conf.int[1], fish_res$conf.int[2]))
cat(sprintf("p-value: %.3g\n", fish_res$p.value))
```

## Figure 2i extended — HFD vs Leptin overlap in AgRP and POMC neurons

```{r panel_hfd_lep_overlap_agrp_pomc}
# Same analysis as Lepr/Glp1r but for AgRP and POMC
make_overlap_plot <- function(ct_diet, ct_lepip, ct_label) {
    # Diet DE
    pb_d <- edger_prep(
        subset(arc_diet_labeled, subset = hash.mcl.ID != "REFED2"),
        celltype_column = "seurat_clusters", trt_group = "treatment", celltypes = ct_diet
    )
    pb_d$samples$treatment <- factor(pb_d$samples$treatment, levels = c("Chow", "HFD", "Refeed", "Fast"))
    design_d <- model.matrix(~treatment + hash_pool, data = pb_d$samples)
    pb_d <- estimateDisp(pb_d, design_d, robust = TRUE)
    fit_d <- glmQLFit(pb_d, design_d, robust = TRUE)
    diet <- topTags(glmQLFTest(fit_d, coef = 2), n = Inf)$table %>%
        dplyr::select(gene, logFC_hfd = logFC, FDR_hfd = FDR)

    # Leptin DE (WT only)
    pb_l <- pseudobulk_lepip[[ct_lepip]]
    pb_l <- pb_l[, pb_l$samples$dataset == "lepip" & pb_l$samples$time != 24 & pb_l$samples$geno == "wt"]
    colnames(pb_l) <- make.unique(colnames(pb_l))
    pb_l$samples$treatment <- factor(pb_l$samples$treatment, levels = c("Sal", "Lep"))
    pb_l$samples$time <- factor(pb_l$samples$time)
    design_l <- model.matrix(~treatment + time + hash_pool, data = pb_l$samples)

    # Check rank
    if (qr(design_l)$rank < ncol(design_l)) {
        design_l <- model.matrix(~treatment + time, data = pb_l$samples)
    }

    fit_l <- voomLmFit(pb_l, design_l)
    fit_l <- eBayes(fit_l, robust = TRUE)
    lep <- topTable(fit_l, coef = "treatmentLep", n = Inf) %>%
        dplyr::select(gene, logFC_lep = logFC, FDR_lep = adj.P.Val)

    # Merge
    ov <- inner_join(diet, lep, by = "gene") %>%
        mutate(same_dir = sign(logFC_hfd) == sign(logFC_lep),
               sig_group = case_when(
                   FDR_hfd < 0.05 & FDR_lep < 0.05 & same_dir ~ "Concordant",
                   FDR_hfd < 0.05 & FDR_lep < 0.05 & !same_dir ~ "Discordant",
                   FDR_hfd < 0.05 ~ "Diet only",
                   FDR_lep < 0.05 ~ "Leptin only",
                   TRUE ~ "Neither"
               ))

    hfd_sig <- ov %>% filter(FDR_hfd < 0.05)
    n_conc <- sum(hfd_sig$sig_group == "Concordant")

    if (nrow(hfd_sig) > 5) {
        ct_r <- cor.test(hfd_sig$logFC_hfd, hfd_sig$logFC_lep)
        r_label <- sprintf("r = %.2f, p = %.2e, n = %d", ct_r$estimate, ct_r$p.value, nrow(hfd_sig))
    } else {
        r_label <- sprintf("n = %d HFD DE genes", nrow(hfd_sig))
    }

    p <- ggplot(hfd_sig, aes(logFC_hfd, logFC_lep)) +
        geom_hline(yintercept = 0, lty = 2, color = "grey50", linewidth = 0.3) +
        geom_vline(xintercept = 0, lty = 2, color = "grey50", linewidth = 0.3) +
        geom_point(data = ~filter(.x, sig_group != "Concordant"), color = "grey60", size = 0.6, alpha = 0.5) +
        geom_point(data = ~filter(.x, sig_group == "Concordant"), color = "#C0392B", size = 0.8, alpha = 0.6) +
        geom_smooth(method = "lm", color = "black", se = FALSE, linewidth = 0.5) +
        annotate("text", x = -Inf, y = Inf, label = r_label, hjust = -0.1, vjust = 1.3, size = 2.5) +
        annotate("text", x = Inf, y = -Inf,
                 label = paste0(n_conc, " concordant / ", nrow(hfd_sig), " HFD DE"),
                 hjust = 1.1, vjust = -0.5, size = 2, color = "grey40") +
        labs(x = "log2FC (HFD vs Chow)", y = "log2FC (Leptin vs Saline)", title = ct_label) +
        cowplot::theme_cowplot(font_size = BASE_SIZE)

    cat(ct_label, ": ", nrow(hfd_sig), " HFD DE, ", n_conc, " concordant\n")
    if (nrow(hfd_sig) > 5) cat("  r = ", round(ct_r$estimate, 3), " p = ", signif(ct_r$p.value, 3), "\n")

    return(p)
}

p_agrp <- make_overlap_plot("Agrp", "Agrp", "AgRP neurons")
p_pomc <- make_overlap_plot("29", "29", "POMC neurons")

# Combine all three
p_combined <- p_overlap + p_agrp + p_pomc +
    patchwork::plot_layout(nrow = 1) &
    theme(legend.position = "none")

ggsave(paste0(fig_path_2, "figure2i_hfd_lep_overlap_all.pdf"), plot = p_combined, h = 4, w = 12)
ggsave(paste0(fig_path_2, "figure2i_hfd_lep_overlap_agrp.pdf"), plot = p_agrp, h = 4, w = 4.5)
ggsave(paste0(fig_path_2, "figure2i_hfd_lep_overlap_pomc.pdf"), plot = p_pomc, h = 4, w = 4.5)
```

## Supplemental — AgRP vs Glp1r/Lepr leptin response overlap

```{r panel_supp_agrp_glp1r_lep_overlap}
# Self-contained: rebuilds both models from scratch

# --- AgRP neurons: WT leptin response ---
pb_a <- pseudobulk_lepip[["Agrp"]]
pb_a <- pb_a[, pb_a$samples$dataset == "lepip" & pb_a$samples$time != 24 & pb_a$samples$geno == "wt"]
colnames(pb_a) <- paste0(colnames(pb_a), "_", pb_a$samples$time)
pb_a$samples$treatment <- factor(pb_a$samples$treatment, levels = c("Sal", "Lep"))
pb_a$samples$time      <- factor(pb_a$samples$time, levels = c("1", "3", "6"))
design_a <- model.matrix(~treatment + time + hash_pool, data = pb_a$samples)
colnames(design_a) <- gsub(":", "_", gsub("[()]", "", colnames(design_a)))
fit2_a   <- eBayes(voomLmFit(pb_a, design_a), robust = TRUE)
agrp_all <- topTable(fit2_a, coef = "treatmentLep", n = Inf) %>%
    dplyr::select(gene, logFC_agrp = logFC, FDR_agrp = adj.P.Val)

# --- Glp1r/Lepr neurons (cluster 12): WT-only leptin response ---
pb_g <- pseudobulk_lepip[["12"]]
pb_g <- pb_g[, pb_g$samples$dataset == "lepip" & pb_g$samples$time != 24 & pb_g$samples$geno == "wt"]
colnames(pb_g) <- paste0(colnames(pb_g), "_", pb_g$samples$time)
pb_g$samples$treatment <- factor(pb_g$samples$treatment, levels = c("Sal", "Lep"))
pb_g$samples$time      <- factor(pb_g$samples$time, levels = c("1", "3", "6"))
design_g <- model.matrix(~treatment + time + hash_pool, data = pb_g$samples)
colnames(design_g) <- gsub(":", "_", gsub("[()]", "", colnames(design_g)))
fit2_g   <- eBayes(voomLmFit(pb_g, design_g), robust = TRUE)
glp_all  <- topTable(fit2_g, coef = "treatmentLep", n = Inf) %>%
    dplyr::select(gene, logFC_glp = logFC, FDR_glp = adj.P.Val)

# --- Merge on shared expressed genes ---
overlap_ag <- inner_join(agrp_all, glp_all, by = "gene") %>%
    mutate(
        same_dir  = sign(logFC_agrp) == sign(logFC_glp),
        sig_group = case_when(
            FDR_agrp < 0.05 & FDR_glp < 0.05 & same_dir  ~ "Concordant",
            FDR_agrp < 0.05 & FDR_glp < 0.05 & !same_dir ~ "Discordant",
            FDR_agrp < 0.05                               ~ "AgRP only",
            FDR_glp  < 0.05                               ~ "Glp1r only",
            TRUE                                           ~ "Neither"
        )
    )

write_csv(overlap_ag, paste0(tab_path_2, "agrp_glp1r_leptin_overlap.csv"))

count_df <- overlap_ag %>%
    filter(sig_group != "Neither") %>%
    dplyr::count(sig_group) %>%
    mutate(sig_group = factor(sig_group, c("AgRP only", "Glp1r only", "Concordant", "Discordant")))

cat("Leptin-regulated DEG overlap (AgRP vs Glp1r/Lepr, FDR < 0.05):\n")
print(count_df)

overlap_cols <- c("Concordant"  = "#C0392B",
                  "Discordant"  = "#8E44AD",
                  "AgRP only"   = "#9A84B7",
                  "Glp1r only"  = "#B85182")

p_bar <- ggplot(count_df, aes(sig_group, n, fill = sig_group)) +
    geom_col(width = 0.7) +
    geom_text(aes(label = n), vjust = -0.4, size = 2.5) +
    scale_fill_manual(values = overlap_cols) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 40, hjust = 1)) +
    labs(x = NULL, y = "Number of DEGs")

ggsave(paste0(fig_path_2, "supp_agrp_glp1r_lep_overlap_bar.pdf"), plot = p_bar, h = 4, w = 4)

cor_ag    <- cor.test(overlap_ag$logFC_agrp, overlap_ag$logFC_glp, method = "pearson")
cor_label <- sprintf("r = %.2f, p = %.2e", cor_ag$estimate, cor_ag$p.value)

p_scatter <- ggplot(overlap_ag, aes(logFC_agrp, logFC_glp, color = sig_group)) +
    geom_point(size = 0.4, alpha = 0.4) +
    geom_smooth(method = "lm", color = "black", se = FALSE, linewidth = 0.5) +
    scale_color_manual(values = c(overlap_cols, "Neither" = "grey80")) +
    annotate("text", x = -Inf, y = Inf, label = cor_label,
             hjust = -0.1, vjust = 1.3, size = 2.5) +
    labs(x = "AgRP logFC (Leptin vs Sal)",
         y = "Glp1r/Lepr logFC (Leptin vs Sal)",
         color = NULL) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(legend.position = "bottom", legend.text = element_text(size = BASE_SIZE - 1))

ggsave(paste0(fig_path_2, "supp_agrp_glp1r_lep_scatter.pdf"), plot = p_scatter, h = 4, w = 4)

cont_ag  <- with(overlap_ag, table(
    AgRP_sig      = FDR_agrp < 0.05,
    Glp1r_concord = FDR_glp  < 0.05 & same_dir
))
fish_ag <- fisher.test(cont_ag)

cat("\nFisher's exact test — concordant leptin response: AgRP vs Glp1r/Lepr\n")
print(cont_ag)
cat(sprintf("Odds ratio: %.2f  (95%% CI: %.2f-%.2f)\n",
            fish_ag$estimate, fish_ag$conf.int[1], fish_ag$conf.int[2]))
cat(sprintf("p-value: %.3g\n", fish_ag$p.value))
```
