---
title: "Figure 4 — Direct leptin action on Lepr/Glp1r neurons determines diet response"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup_libraries_data}
library(Seurat)
library(tidyverse)
library(targets)
library(edgeR)
library(limma)
library(lme4)
library(emmeans)
library(ggrepel)
library(patchwork)
library(magrittr)
library(MASS)
library(janitor)

source(here::here("R/basic_processing.R"))
source(here::here("R/analysis_wrappers.R"))
source(here::here("R/helper_functions.R"))

fig_path_4 <- here::here("output_figures/paper/figure4/")
tab_path_4 <- here::here("output_figures/paper/figure4/tables/")
dir.create(fig_path_4, recursive = TRUE, showWarnings = FALSE)
dir.create(tab_path_4, recursive = TRUE, showWarnings = FALSE)

# Global base font size — change here to update all panels
BASE_SIZE <- 8
geno_cols  <- c("mut" = "#C0392B", "wt" = "grey70")

# Leptin gene signature
# LGS genes derived from MINT sPLS-DA (see paper/lgs.Rmd)
lgs_genes <- read.csv(here::here("output_figures/paper/lgs/lgs_genes.csv"))
leptin_geneset <- lgs_genes$gene[lgs_genes$value.var < 0]

# Load only what Figure 4 core panels need
targets::tar_config_set(store = here::here("_targets"))
targets::tar_load(arc_diet_labeled)
targets::tar_load(arc_glp1rko_labeled)
targets::tar_load(pb_glp1rko)
targets::tar_load(celltypes_glp1rko)
names(pb_glp1rko) <- celltypes_glp1rko
```

## Identify Lepr-expressing clusters from diet reference

```{r lepr_clusters}
DefaultAssay(arc_diet_labeled) <- "RNA"
lepr_exp <- AverageExpression(arc_diet_labeled, features = "Lepr", group.by = "seurat_clusters")
lepclus  <- names(as.data.frame(lepr_exp$RNA))[which(lepr_exp$RNA > 1)]
lepclus  <- lepclus[-5]   # remove low-quality cluster, consistent with figure2
```

## Figure 4e — LGS across Lepr+ cell types: GLP1R/Lepr-KO

```{r panel_4e_lgs_glp1rko}
# Compute LGS scores in GLP1R-KO dataset
DefaultAssay(arc_glp1rko_labeled) <- "RNA"
arc_glp1rko_labeled@meta.data <- arc_glp1rko_labeled@meta.data[,
    !grepl("Cluster", colnames(arc_glp1rko_labeled@meta.data))]
arc_glp1rko_labeled <- AddModuleScore(arc_glp1rko_labeled, features = list(leptin_geneset))

# Pseudobulk LGS per animal × cell type — all Lepr+ clusters
gko_pb <- arc_glp1rko_labeled[[]] %>%
    dplyr::select(geno, diet, predicted.celltype, hash.mcl.ID, hash_pool, sex, starts_with("Cluster")) %>%
    filter(predicted.celltype %in% lepclus) %>%
    group_by(geno, diet, predicted.celltype, hash.mcl.ID, hash_pool, sex) %>%
    summarise(mean_lgs = mean(Cluster1), .groups = "drop") %>%
    mutate(geno = factor(geno, c("wt", "mut")),
           diet = factor(diet, c("Chow", "DIO")))

# Per-cell-type model: WT - KO emmeans contrasts by diet
gko_res <- gko_pb %>%
    nest(data = -predicted.celltype) %>%
    mutate(mod = purrr::map(data, ~lm(mean_lgs ~ geno * diet + sex, data = .x)),
           emm = purrr::map(mod, ~emmeans::emmeans(.x, pairwise ~ geno | diet)$contrasts %>% data.frame())) %>%
    dplyr::select(predicted.celltype, emm) %>%
    unnest(emm) %>%
    mutate(padj = p.adjust(p.value, "BH")) %>%
    dplyr::rename(ct = predicted.celltype)

write_csv(gko_res, paste0(tab_path_4, "lgs_glp1rko_per_cluster_stats.csv"))

# Overall model on Lepr/Glp1r neurons (cluster 12) — report beta and interaction p
clus12_pb    <- gko_pb %>% filter(predicted.celltype == "12")
mod_overall  <- lm(mean_lgs ~ geno * diet + sex, data = clus12_pb)
beta_geno    <- coef(summary(mod_overall))["genomut", "Estimate"]
p_geno       <- coef(summary(mod_overall))["genomut", "Pr(>|t|)"]
p_interaction <- summary(mod_overall)$coefficients["genomut:dietDIO", "Pr(>|t|)"]
cat(sprintf("Lepr/Glp1r neurons — geno effect: beta=%.3f, p=%.2e\n", beta_geno, p_geno))
cat(sprintf("Genotype x diet interaction: p=%.2e\n", p_interaction))

p_lgs_glp1rko_clusters <- gko_res %>%
    ggplot() +
    geom_errorbar(aes(y = ct, x = estimate,
                      xmin = estimate - SE * 1.96, xmax = estimate + SE * 1.96,
                      color = diet), width = 0, position = position_dodge(width = 0.5), linewidth = 0.6) +
    geom_point(shape = 21, stroke = 0.8, size = 2.5,
               aes(y = ct, x = estimate, fill = diet,
                   color = I(ifelse(padj < 0.05, "black", "grey60"))),
               position = position_dodge(width = 0.5)) +
    geom_vline(xintercept = 0, lty = 2, linewidth = 0.4, color = "grey40") +
    scale_fill_manual(values  = c("DIO" = "#E8682F", "Chow" = "#F0AD74")) +
    scale_color_manual(values = c("DIO" = "#E8682F", "Chow" = "#F0AD74")) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    cowplot::panel_border() +
    labs(x = "WT - KO LGS (estimated difference)", y = NULL, fill = "Diet", color = "Diet")

ggsave(paste0(fig_path_4, "figure4e_lgs_glp1rko_clusters.pdf"),
       plot = p_lgs_glp1rko_clusters, h = 4, w = 4)
```

## Figure 4f — PCA: GLP1R/Lepr-KO cluster 12 projected onto diet PCA space

```{r panel_4f_pca_glp1rko}
# Project Lepr/Glp1r (cluster 12) KO cells onto reference diet PCA
ref_data <- subset(arc_diet_labeled, subset = treatment %in% c("HFD", "Chow", "Fast"))
celltype <- "12"

ref_subset   <- subset(ref_data, subset = seurat_clusters == celltype)
new_subset   <- subset(arc_glp1rko_labeled, subset = predicted.celltype == celltype)
shared_feats <- intersect(rownames(ref_subset), rownames(new_subset))

ref_subset <- process_seurat(subset(ref_subset, features = shared_feats),
                              method = "integrate", batch = "hash_pool",
                              cluster = FALSE, nfeats = 5000)
ref_mat       <- ref_subset@assays$integrated@scale.data
ref_pca       <- prcomp(t(ref_mat), center = TRUE, scale. = FALSE)

new_subset    <- NormalizeData(new_subset) %>% ScaleData(features = VariableFeatures(ref_subset))
scale_mat     <- new_subset@assays$RNA@scale.data
scale_mat     <- scale_mat[match(rownames(ref_mat), rownames(scale_mat)), ]
new_pca_scores <- as.matrix(t(scale_mat)) %*% ref_pca$rotation %>% data.frame()

diet_pca_cols <- c("HFD" = "#E8682F", "Chow" = "#F0AD74", "Fast" = "#165C5B")
geno_cols <- c("Control" = "#4A90D9", "Glp1rLepr-KO" = "#C0392B")

# Helper: plot DIO KO vs WT on colored reference background + permutation test
plot_pca_dio_shift <- function(ref_bg_df, ko_pts_df, title_label, n_perm = 1000) {
    ko_dio <- ko_pts_df %>% filter(diet == "HFD")

    # Permutation test: is the PC1 shift between genotypes significant?
    obs_diff <- abs(diff(tapply(ko_dio$PC1, ko_dio$geno, mean)))
    perm_diffs <- replicate(n_perm, {
        shuf <- sample(ko_dio$geno)
        abs(diff(tapply(ko_dio$PC1, shuf, mean)))
    })
    perm_p <- (sum(perm_diffs >= obs_diff) + 1) / (n_perm + 1)
    cat(title_label, "- PC1 shift permutation test: obs_diff =", round(obs_diff, 3),
        ", p =", round(perm_p, 4), "\n")

    # Compute centroids for Fast and HFD reference groups
    ref_centroids <- ref_bg_df %>%
        filter(diet %in% c("Fast", "HFD")) %>%
        group_by(diet) %>%
        summarise(PC1 = mean(PC1), PC2 = mean(PC2), .groups = "drop")

    p <- ggplot() +
        geom_point(data = ref_bg_df, aes(PC1, PC2, color = diet), size = 0.3, alpha = 0.15) +
        scale_color_manual(values = diet_pca_cols, guide = guide_legend(override.aes = list(size = 2, alpha = 1))) +
        ggnewscale::new_scale_color() +
        geom_point(data = ko_dio, aes(PC1, PC2, color = geno), size = 0.8, alpha = 0.6) +
        scale_color_manual(values = geno_cols, guide = guide_legend(override.aes = list(size = 2, alpha = 1))) +
        annotate("point", x = ref_centroids$PC1, y = ref_centroids$PC2,
                 shape = 4, size = 5, stroke = 2, color = diet_pca_cols[ref_centroids$diet]) +
        cowplot::theme_cowplot(font_size = BASE_SIZE) +
        labs(x = "PC1", y = "PC2", title = title_label)

    return(list(plot = p, perm_p = perm_p, obs_diff = obs_diff))
}

# Reference background and KO points for Lepr/Glp1r
ref_bg <- data.frame(ref_pca$x[, 1:2], diet = ref_subset@meta.data[["treatment"]])
ko_pts <- new_pca_scores[, 1:2] %>%
    bind_cols(new_subset[[]] %>% dplyr::select(geno, diet)) %>%
    mutate(diet = ifelse(diet == "DIO", "HFD", diet),
           geno = factor(geno, c("wt", "mut"),
                         labels = c("Control", "Glp1rLepr-KO")))

set.seed(42)
res_glp1r <- plot_pca_dio_shift(ref_bg, ko_pts, "Lepr/Glp1r neurons (DIO)")

ggsave(paste0(fig_path_4, "figure4f_pca_glp1rko.pdf"), plot = res_glp1r$plot, h = 4, w = 5)
```

## Figure 4f2 — PCA: GLP1R/Lepr-KO AgRP neurons projected onto diet PCA space

```{r panel_4f2_pca_agrp}
# Project AgRP KO cells onto reference diet PCA (same approach as Lepr/Glp1r above)
celltype_agrp <- "Agrp"

ref_subset_agrp <- subset(ref_data, subset = seurat_clusters == celltype_agrp)
new_subset_agrp <- subset(arc_glp1rko_labeled, subset = predicted.celltype == celltype_agrp)
shared_feats_agrp <- intersect(rownames(ref_subset_agrp), rownames(new_subset_agrp))

ref_subset_agrp <- process_seurat(subset(ref_subset_agrp, features = shared_feats_agrp),
                                   method = "integrate", batch = "hash_pool",
                                   cluster = FALSE, nfeats = 5000)
ref_mat_agrp       <- ref_subset_agrp@assays$integrated@scale.data
ref_pca_agrp       <- prcomp(t(ref_mat_agrp), center = TRUE, scale. = FALSE)

new_subset_agrp    <- NormalizeData(new_subset_agrp) %>% ScaleData(features = VariableFeatures(ref_subset_agrp))
scale_mat_agrp     <- new_subset_agrp@assays$RNA@scale.data
scale_mat_agrp     <- scale_mat_agrp[match(rownames(ref_mat_agrp), rownames(scale_mat_agrp)), ]
new_pca_agrp       <- as.matrix(t(scale_mat_agrp)) %*% ref_pca_agrp$rotation %>% data.frame()

ref_bg_agrp <- data.frame(ref_pca_agrp$x[, 1:2], diet = ref_subset_agrp@meta.data[["treatment"]])
ko_pts_agrp <- new_pca_agrp[, 1:2] %>%
    bind_cols(new_subset_agrp[[]] %>% dplyr::select(geno, diet)) %>%
    mutate(diet = ifelse(diet == "DIO", "HFD", diet),
           geno = factor(geno, c("wt", "mut"),
                         labels = c("Control", "Glp1rLepr-KO")))

res_agrp <- plot_pca_dio_shift(ref_bg_agrp, ko_pts_agrp, "AgRP neurons (DIO)")

ggsave(paste0(fig_path_4, "figure4f2_pca_agrp.pdf"), plot = res_agrp$plot, h = 4, w = 5)
```

## Figure 4g–h — DE volcano: GLP1R/Lepr-KO vs WT in Lepr/Glp1r neurons (Chow and DIO)

```{r panel_4g_h_volcano_glp1rko}
pbd <- pb_glp1rko[["12"]]
pbd$samples$group <- interaction(pbd$samples$geno, pbd$samples$diet)
design <- model.matrix(~ 0 + group + sex, data = pbd$samples)
colnames(design) <- gsub(":", "_", colnames(design))

fitd <- voomLmFit(pbd, design)
contrast.matrix <- makeContrasts(
    DIO  = groupmut.DIO  - groupwt.DIO,
    Chow = groupmut.Chow - groupwt.Chow,
    levels = design
)
fit2 <- contrasts.fit(fitd, contrast.matrix)
fit2 <- eBayes(fit2, robust = FALSE)

glp12_chow_degs <- topTable(fit2, n = Inf, coef = "Chow") %>% mutate(sig = adj.P.Val < 0.05)
glp12_dio_degs  <- topTable(fit2, n = Inf, coef = "DIO")  %>% mutate(sig = adj.P.Val < 0.05)

cat("Chow DEGs (FDR<0.05):", sum(glp12_chow_degs$sig), "\n")
cat("DIO  DEGs (FDR<0.05):", sum(glp12_dio_degs$sig),  "\n")

write_csv(glp12_chow_degs, paste0(tab_path_4, "glp1rko_clus12_chow_degs.csv"))
write_csv(glp12_dio_degs,  paste0(tab_path_4, "glp1rko_clus12_dio_degs.csv"))

# Genes the manuscript explicitly names on the DIO volcano
force_label_dio <- c("Fos", "Vgf", "Homer1",
                     "Gabra3", "Gabra4", "Gabra5", "Gabrb1", "Gabrb2", "Gabrb3")

volc_glp12_dat <- bind_rows(
    glp12_chow_degs %>% mutate(comparison = "Chow"),
    glp12_dio_degs  %>% mutate(comparison = "DIO")
)

# Top-10 by FDR per panel, plus forced manuscript genes on DIO panel
volc_glp12_labs <- bind_rows(
    volc_glp12_dat %>% filter(sig) %>%
        group_by(comparison) %>% slice_min(adj.P.Val, n = 10) %>% ungroup(),
    volc_glp12_dat %>% filter(comparison == "DIO", gene %in% force_label_dio)
) %>% distinct(comparison, gene, .keep_all = TRUE)

p_volc_glp12 <- volc_glp12_dat %>%
    ggplot(aes(logFC, -log10(adj.P.Val))) +
    geom_point(aes(color = I(ifelse(sig, "#2C3E50", "grey80"))), alpha=0.5, shape = 16, size = 0.6) +
    ggrepel::geom_text_repel(data = volc_glp12_labs, aes(label = gene), color="red",fontface = "bold",
                              size = 3, max.overlaps = 15, min.segment.length = 0,
                              segment.size = 0.3, show.legend = FALSE) +
    geom_hline(yintercept = -log10(0.05), lty = 2, linewidth = 0.3, color = "grey50") +
    facet_wrap(~ comparison, nrow = 1, scales="free_x") +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    cowplot::panel_border() +
    theme(strip.background = element_blank()) +
    labs(x = "log2FC (KO - WT)", y = "-log10(adj. p-value)")

ggsave(paste0(fig_path_4, "figure4g_h_volcano_glp1rko_clus12.pdf"), plot = p_volc_glp12, h = 5, w = 12)
```

## Figure 4i — AgRP neurons in DIO KO adopt fasting-like transcriptional state

```{r panel_4i_volcano_agrp_fasting}
# Diet-regulated genes in AgRP neurons — split fasting into directional sets
# comp: "hvc" = HFD vs Chow, "fvc" = Fast vs Chow
diet_pairwise <- read.csv(here::here("results/diet_pairwise_deg.csv"))

fast_up   <- diet_pairwise %>%
    filter(ct == "Agrp", comp == "fvc", adj.P.Val < 0.05, logFC > 1) %>% pull(gene)
fast_down <- diet_pairwise %>%
    filter(ct == "Agrp", comp == "fvc", adj.P.Val < 0.05, logFC < -1) %>% pull(gene)
hfd_sig   <- diet_pairwise %>%
    filter(ct == "Agrp", comp == "hvc", adj.P.Val < 0.05) %>% pull(gene)

cat("Fasting-upregulated AgRP genes (|logFC|>1):", length(fast_up), "\n")
cat("Fasting-downregulated AgRP genes (|logFC|>1):", length(fast_down), "\n")
cat("HFD-regulated AgRP genes:", length(hfd_sig), "\n")

# GLP1R/Lepr-KO vs WT in AgRP neurons — DIO only
pbd_av <- pb_glp1rko[["Agrp"]]
pbd_av$samples$group <- interaction(pbd_av$samples$geno, pbd_av$samples$diet)
design_av <- model.matrix(~ 0 + group + sex, data = pbd_av$samples)
colnames(design_av) <- gsub(":", "_", colnames(design_av))

fit2_av <- contrasts.fit(
    voomLmFit(pbd_av, design_av),
    makeContrasts(DIO = groupmut.DIO - groupwt.DIO, levels = design_av)
) %>% eBayes(robust = FALSE)

dio_av <- topTable(fit2_av, n = Inf, coef = "DIO") %>%
    mutate(
        # Only highlight fasting-regulated genes; HFD genes uncolored
        diet_group = case_when(
            gene %in% fast_up   ~ "Fasting up",
            gene %in% fast_down ~ "Fasting dn",
            TRUE                ~ "Neither"
        ),
        diet_group = factor(diet_group, c("Fasting up", "Fasting dn", "Neither"))
    )

write_csv(dio_av, paste0(tab_path_4, "glp1rko_agrp_dio_diet_annotated.csv"))
cat("KO DIO genes by diet annotation:\n")
print(count(dio_av, diet_group))

diet_volc_cols <- c("Fasting up" = "#E07B39", "Fasting dn" = "#2E86AB", "Neither" = "grey80")

# Label only KO-significant fasting-concordant genes — HFD genes NOT labeled
diet_volc_labs <- dio_av %>%
    filter(adj.P.Val < 0.05) %>%
    filter((logFC > 0 & diet_group == "Fasting up") |
           (logFC < 0 & diet_group == "Fasting dn"))

p_volc_diet <- dio_av %>%
    arrange(desc(diet_group == "Neither")) %>%
    ggplot(aes(logFC, -log10(adj.P.Val), color = diet_group)) +
    geom_point(aes(size  = diet_group != "Neither",
                   alpha = diet_group != "Neither"), shape = 16) +
    scale_size_manual(values  = c("TRUE" = 1.2, "FALSE" = 0.4), guide = "none") +
    scale_alpha_manual(values = c("TRUE" = 0.85, "FALSE" = 0.2), guide = "none") +
    scale_color_manual(values = diet_volc_cols) +
    ggrepel::geom_text_repel(data = diet_volc_labs, aes(label = ifelse(adj.P.Val<5e-3, gene, NA_character_)),
                              size = 2, max.overlaps = 15, min.segment.length = 0,
                              segment.size = 0.3, show.legend = FALSE) +
    geom_hline(yintercept = -log10(0.05), lty = 2, color = "grey50", linewidth = 0.3) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(legend.position = "right") +
    labs(x = "log2FC (KO - WT, DIO)", y = "-log10(adj. p-value)",
         color = "Fasting response\n(AgRP neurons)")

ggsave(paste0(fig_path_4, "figure4i_volcano_agrp_fasting.pdf"), plot = p_volc_diet, h = 5, w = 4)

# --- Directional enrichment: KO in DIO recapitulates the fasting transcriptome ---
ko_up   <- dio_av %>% filter(adj.P.Val < 0.05, logFC > 0) %>% pull(gene)
ko_down <- dio_av %>% filter(adj.P.Val < 0.05, logFC < 0) %>% pull(gene)
bg      <- dio_av$gene

ft1 <- fisher.test(
    matrix(c(sum(bg %in% ko_up  &  bg %in% fast_up),
              sum(bg %in% ko_up  & !bg %in% fast_up),
              sum(!bg %in% ko_up &  bg %in% fast_up),
              sum(!bg %in% ko_up & !bg %in% fast_up)), nrow = 2),
    alternative = "greater")
cat("Fasting up intersect KO-upregulated (DIO):\n")
cat("  n =", sum(bg %in% ko_up & bg %in% fast_up),
    " OR =", round(ft1$estimate, 2), " p =", signif(ft1$p.value, 3), "\n\n")

ft2 <- fisher.test(
    matrix(c(sum(bg %in% ko_down  &  bg %in% fast_down),
              sum(bg %in% ko_down  & !bg %in% fast_down),
              sum(!bg %in% ko_down &  bg %in% fast_down),
              sum(!bg %in% ko_down & !bg %in% fast_down)), nrow = 2),
    alternative = "greater")
cat("Fasting dn intersect KO-downregulated (DIO):\n")
cat("  n =", sum(bg %in% ko_down & bg %in% fast_down),
    " OR =", round(ft2$estimate, 2), " p =", signif(ft2$p.value, 3), "\n")

# DIO-only volcano for Lepr/Glp1r (extract from full data)
p_volc_glp12_dio <- volc_glp12_dat %>%
    filter(comparison == "DIO") %>%
    ggplot(aes(logFC, -log10(adj.P.Val))) +
    geom_point(aes(color = I(ifelse(sig, "#2C3E50", "grey80"))), alpha = 0.5, shape = 16, size = 0.6) +
    ggrepel::geom_text_repel(
        data = volc_glp12_labs %>% filter(comparison == "DIO"),
        aes(label = gene), color = "red", fontface = "bold",
        size = 2, max.overlaps = 15, min.segment.length = 0,
        segment.size = 0.3, show.legend = FALSE) +
    geom_hline(yintercept = -log10(0.05), lty = 2, linewidth = 0.3, color = "grey50") +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    labs(x = "log2FC (KO - WT, DIO)", y = "-log10(adj. p-value)", title = "Lepr/Glp1r neurons (DIO)")

# Composite: 2 rows × 2 cols
# Top:    Lepr/Glp1r PCA  |  Lepr/Glp1r DIO volcano
# Bottom: AgRP PCA        |  AgRP DIO fasting volcano
p_composite <- (res_glp1r$plot + p_volc_glp12_dio) /
               (res_agrp$plot + p_volc_diet) +
    patchwork::plot_annotation(tag_levels = "a") &
    theme(plot.tag = element_text(face = "bold", size = 10))

ggsave(paste0(fig_path_4, "figure4_pca_volcano_composite.pdf"),
       plot = p_composite, h = 5, w = 14)

# Save individual plot objects
saveRDS(p_lgs_glp1rko_clusters, paste0(fig_path_4, "p_fig4e.rds"))
saveRDS(res_glp1r$plot,          paste0(fig_path_4, "p_fig4f.rds"))
saveRDS(p_volc_glp12,           paste0(fig_path_4, "p_fig4gh.rds"))
saveRDS(p_volc_diet,            paste0(fig_path_4, "p_fig4i.rds"))
```
