---
title: "Reviewer Response — mTOR pathway analysis in Lepr/Glp1r neurons"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

This analysis addresses a reviewer comment asking whether Lepr/Glp1r neurons
have relatively lower mTOR pathway activity compared to other Lepr+ cell types
in DIO, which could explain their preserved leptin sensitivity. Motivated by
Tan et al. (2025, Cell Metabolism), who showed that increased mTOR activity in
POMC neurons drives leptin resistance in DIO mice.

```{r libraries_and_data}
library(Seurat)
library(tidyverse)
library(emmeans)
library(patchwork)
library(ComplexHeatmap)
library(circlize)

source(here::here("R/basic_processing.R"))
source(here::here("R/analysis_wrappers.R"))
source(here::here("R/helper_functions.R"))

BASE_SIZE <- 8
fig_path <- here::here("output_figures/paper/reviewer_mtor/")
tab_path <- here::here("output_figures/paper/reviewer_mtor/tables/")
dir.create(fig_path, recursive = TRUE, showWarnings = FALSE)
dir.create(tab_path, recursive = TRUE, showWarnings = FALSE)

diet_cols <- c("HFD" = "#E8682F", "Chow" = "#F0AD74",
               "Refeed" = "#58C2AF", "Fast" = "#165C5B")

# Load targets
targets::tar_config_set(store = here::here("_targets"))
targets::tar_load(arc_diet_labeled)
targets::tar_load(pb_alldiet)
targets::tar_load(celltypes_diet)
names(pb_alldiet) <- celltypes_diet

# Lepr clusters
DefaultAssay(arc_diet_labeled) <- "RNA"
all_lepr_clusters <- c("Agrp", "12", "29", "26", "28")
all_lepr_labels   <- c("Agrp" = "AgRP", "12" = "Lepr/Glp1r", "29" = "POMC",
                        "26" = "Tbx19", "28" = "Irx3")
```

## Define mTOR pathway gene sets

```{r mtor_genesets}
# Core mTORC1 signaling — genes from Reactome + key regulators
# Positive regulators of mTOR (higher expression = more mTOR activity)
mtor_positive <- c("Mtor", "Rptor", "Mlst8", "Rheb", "Rraga", "Rragb",
                    "Rragc", "Rragd", "Lamtor1", "Lamtor2", "Lamtor3",
                    "Lamtor4", "Lamtor5", "Akt1", "Akt2", "Akt3",
                    "Pik3ca", "Pik3cb", "Pik3r1")

# Negative regulators of mTOR (higher expression = less mTOR activity)
mtor_negative <- c("Tsc1", "Tsc2", "Pten", "Ddit4", "Deptor",
                    "Sesn1", "Sesn2", "Sesn3",
                    "Prkaa1", "Prkaa2", "Stk11")

# mTOR downstream targets / readouts of activity
mtor_downstream <- c("Rps6kb1", "Rps6kb2", "Eif4ebp1", "Eif4ebp2",
                      "Eif4e", "Eif4b", "Rps6")

# Full mTOR module for AddModuleScore
mtor_all <- c(mtor_positive, mtor_negative, mtor_downstream)

# Check which genes are present in the dataset
genes_in_data <- intersect(mtor_all, rownames(arc_diet_labeled))
missing <- setdiff(mtor_all, rownames(arc_diet_labeled))
cat("mTOR genes found:", length(genes_in_data), "/", length(mtor_all), "\n")
cat("Missing:", paste(missing, collapse = ", "), "\n")

mtor_positive_present   <- intersect(mtor_positive, rownames(arc_diet_labeled))
mtor_negative_present   <- intersect(mtor_negative, rownames(arc_diet_labeled))
mtor_downstream_present <- intersect(mtor_downstream, rownames(arc_diet_labeled))

cat("\nPositive regulators:", length(mtor_positive_present), "—",
    paste(mtor_positive_present, collapse = ", "), "\n")
cat("Negative regulators:", length(mtor_negative_present), "—",
    paste(mtor_negative_present, collapse = ", "), "\n")
cat("Downstream targets:", length(mtor_downstream_present), "—",
    paste(mtor_downstream_present, collapse = ", "), "\n")
```

## mTOR module score across Lepr clusters and diet conditions

```{r mtor_module_score}
# Score mTOR positive regulators (proxy for mTOR activity)
arc_diet_labeled <- AddModuleScore(
    arc_diet_labeled,
    features = list(mtor_positive_present),
    name = "mTOR_pos_",
    assay = "RNA"
)

# Score mTOR negative regulators
arc_diet_labeled <- AddModuleScore(
    arc_diet_labeled,
    features = list(mtor_negative_present),
    name = "mTOR_neg_",
    assay = "RNA"
)

# Score downstream targets
arc_diet_labeled <- AddModuleScore(
    arc_diet_labeled,
    features = list(mtor_downstream_present),
    name = "mTOR_down_",
    assay = "RNA"
)

# Net mTOR activity estimate: positive regulators - negative regulators
arc_diet_labeled$mTOR_net <- arc_diet_labeled$mTOR_pos_1 - arc_diet_labeled$mTOR_neg_1
```

## Panel A — mTOR positive regulator score: Chow vs HFD across Lepr clusters

```{r panel_a_mtor_pos_boxplot, fig.width=6, fig.height=3.5}
mtor_df <- arc_diet_labeled[[]] %>%
    filter(seurat_clusters %in% all_lepr_clusters,
           treatment %in% c("Chow", "HFD")) %>%
    mutate(ct_label  = all_lepr_labels[as.character(seurat_clusters)],
           ct_label  = factor(ct_label, c("AgRP", "Lepr/Glp1r", "POMC", "Tbx19", "Irx3")),
           treatment = factor(treatment, c("Chow", "HFD"))) %>%
    group_by(hash.mcl.ID, hash_pool, treatment, ct_label) %>%
    summarise(
        mean_mtor_pos  = mean(mTOR_pos_1),
        mean_mtor_neg  = mean(mTOR_neg_1),
        mean_mtor_down = mean(mTOR_down_1),
        mean_mtor_net  = mean(mTOR_net),
        .groups = "drop"
    )

# Stats: per-cluster Chow vs HFD comparison
mtor_stats <- mtor_df %>%
    nest(data = -ct_label) %>%
    mutate(
        mod_pos  = map(data, ~lm(mean_mtor_pos ~ treatment, data = .x)),
        mod_neg  = map(data, ~lm(mean_mtor_neg ~ treatment, data = .x)),
        mod_net  = map(data, ~lm(mean_mtor_net ~ treatment, data = .x)),
        emm_pos  = map(mod_pos,  ~emmeans(.x, pairwise ~ treatment)$contrasts %>% as.data.frame()),
        emm_neg  = map(mod_neg,  ~emmeans(.x, pairwise ~ treatment)$contrasts %>% as.data.frame()),
        emm_net  = map(mod_net,  ~emmeans(.x, pairwise ~ treatment)$contrasts %>% as.data.frame())
    )

stats_pos <- mtor_stats %>% unnest(emm_pos) %>%
    dplyr::select(ct_label, contrast, estimate, SE, p.value) %>%
    mutate(score = "mTOR_positive_regulators")
stats_neg <- mtor_stats %>% unnest(emm_neg) %>%
    dplyr::select(ct_label, contrast, estimate, SE, p.value) %>%
    mutate(score = "mTOR_negative_regulators")
stats_net <- mtor_stats %>% unnest(emm_net) %>%
    dplyr::select(ct_label, contrast, estimate, SE, p.value) %>%
    mutate(score = "mTOR_net_activity")

mtor_all_stats <- bind_rows(stats_pos, stats_neg, stats_net) %>%
    mutate(padj = p.adjust(p.value, "BH"))

write_csv(mtor_all_stats, paste0(tab_path, "mtor_module_chow_vs_hfd_stats.csv"))
cat("\n=== mTOR module score: Chow vs HFD per cluster ===\n")
mtor_all_stats %>%
    mutate(sig = case_when(padj < 0.001 ~ "***", padj < 0.01 ~ "**",
                           padj < 0.05 ~ "*", padj < 0.1 ~ "+", TRUE ~ "ns")) %>%
    print(n = 30)

# Also test cluster × diet interaction (is cluster 12 different from others?)
# Use mixed model: animals contribute observations to multiple clusters
mtor_interaction <- mtor_df %>%
    mutate(is_leprglp1r = ifelse(ct_label == "Lepr/Glp1r", "Lepr/Glp1r", "Other")) %>%
    lme4::lmer(mean_mtor_pos ~ is_leprglp1r * treatment + (1 | hash.mcl.ID), data = .)

cat("\n=== Interaction: Lepr/Glp1r vs Others × Diet (mTOR positive regulators) ===\n")
print(summary(mtor_interaction)$coefficients)
cat("\nType II Anova (Kenward-Roger):\n")
print(car::Anova(mtor_interaction, type = "II", test.statistic = "F"))

# Plot: mTOR positive regulator score
sig_label <- function(p) case_when(p < 0.001 ~ "***", p < 0.01 ~ "**",
                                    p < 0.05 ~ "*", p < 0.1 ~ "+", TRUE ~ "ns")

signif_pos <- stats_pos %>%
    left_join(mtor_all_stats %>% filter(score == "mTOR_positive_regulators") %>%
                  dplyr::select(ct_label, padj), by = "ct_label") %>%
    mutate(label = sig_label(padj)) %>%
    filter(label != "ns")

p_mtor_pos <- ggplot(mtor_df, aes(treatment, mean_mtor_pos, fill = treatment)) +
    geom_boxplot(outlier.shape = NA, width = 0.6, linewidth = 0.35) +
    geom_jitter(width = 0.15, size = 0.8, alpha = 0.6, shape = 16) +
    facet_wrap(~ct_label, nrow = 1, scales = "free_y") +
    scale_fill_manual(values = diet_cols) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.background = element_blank()) +
    labs(x = NULL, y = "mTOR Positive Regulator Score",
         title = "mTOR positive regulators across Lepr+ clusters")

ggsave(paste0(fig_path, "mtor_positive_regulators_chow_hfd.pdf"),
       plot = p_mtor_pos, h = 3.5, w = 6)

# Plot: mTOR net activity
p_mtor_net <- ggplot(mtor_df, aes(treatment, mean_mtor_net, fill = treatment)) +
    geom_boxplot(outlier.shape = NA, width = 0.6, linewidth = 0.35) +
    geom_jitter(width = 0.15, size = 0.8, alpha = 0.6, shape = 16) +
    facet_wrap(~ct_label, nrow = 1, scales = "free_y") +
    scale_fill_manual(values = diet_cols) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.background = element_blank()) +
    labs(x = NULL, y = "Net mTOR Activity Score\n(positive - negative regulators)",
         title = "Net mTOR activity across Lepr+ clusters")

ggsave(paste0(fig_path, "mtor_net_activity_chow_hfd.pdf"),
       plot = p_mtor_net, h = 3.5, w = 6)
```

## Panel B — mTOR module across all 4 diet states

```{r panel_b_mtor_all_diets, fig.width=8, fig.height=3.5}
mtor_alldiets <- arc_diet_labeled[[]] %>%
    filter(seurat_clusters %in% all_lepr_clusters) %>%
    mutate(ct_label  = all_lepr_labels[as.character(seurat_clusters)],
           ct_label  = factor(ct_label, c("AgRP", "Lepr/Glp1r", "POMC", "Tbx19", "Irx3")),
           treatment = fct_relevel(treatment, c("Fast", "Refeed", "Chow", "HFD"))) %>%
    group_by(hash.mcl.ID, hash_pool, treatment, ct_label) %>%
    summarise(mean_mtor_pos = mean(mTOR_pos_1),
              mean_mtor_net = mean(mTOR_net),
              .groups = "drop")

# Stats: all pairwise within each cluster
mtor_alldiets_stats <- mtor_alldiets %>%
    nest(data = -ct_label) %>%
    mutate(
        mod = map(data, ~lm(mean_mtor_pos ~ treatment, data = .x)),
        emm = map(mod, ~emmeans(.x, pairwise ~ treatment)$contrasts %>% as.data.frame())
    ) %>%
    unnest(emm) %>%
    ungroup() %>%
    mutate(padj = p.adjust(p.value, "BH")) %>%
    dplyr::select(ct_label, contrast, estimate, SE, p.value, padj)

write_csv(mtor_alldiets_stats, paste0(tab_path, "mtor_module_all_diets_stats.csv"))

p_mtor_alldiets <- ggplot(mtor_alldiets, aes(treatment, mean_mtor_pos, fill = treatment)) +
    geom_boxplot(outlier.shape = NA, width = 0.6, linewidth = 0.35) +
    geom_jitter(width = 0.15, size = 0.5, alpha = 0.5, shape = 16) +
    facet_wrap(~ct_label, nrow = 1, scales = "free_y") +
    scale_fill_manual(values = diet_cols) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.background = element_blank()) +
    labs(x = NULL, y = "mTOR Positive Regulator Score")

ggsave(paste0(fig_path, "mtor_positive_all_diets.pdf"),
       plot = p_mtor_alldiets, h = 3.5, w = 8)
```

## Panel C — Individual mTOR gene expression heatmap (pseudobulk)

```{r panel_c_heatmap, fig.width=7, fig.height=6}
library(edgeR)

# Build pseudobulk for all Lepr clusters — Chow vs HFD
mtor_genes_present <- unique(c(mtor_positive_present, mtor_negative_present,
                                mtor_downstream_present))

pb_heatmap_data <- list()

for (ct in all_lepr_clusters) {
    # edger_prep already filters genes and normalizes lib sizes internally
    pb <- edger_prep(
        subset(arc_diet_labeled, subset = treatment %in% c("Chow", "HFD")),
        celltype_column = "seurat_clusters", trt_group = "treatment", celltypes = ct
    )

    # logCPM per sample
    lcpm <- cpm(pb, log = TRUE, prior.count = 1)
    genes_avail <- intersect(mtor_genes_present, rownames(lcpm))

    # Mean logCPM per treatment
    chow_samples <- pb$samples$treatment == "Chow"
    hfd_samples  <- pb$samples$treatment == "HFD"

    pb_heatmap_data[[ct]] <- data.frame(
        gene = genes_avail,
        cluster = all_lepr_labels[ct],
        logCPM_Chow = rowMeans(lcpm[genes_avail, chow_samples, drop = FALSE]),
        logCPM_HFD  = rowMeans(lcpm[genes_avail, hfd_samples, drop = FALSE]),
        logFC_HFD   = rowMeans(lcpm[genes_avail, hfd_samples, drop = FALSE]) -
                       rowMeans(lcpm[genes_avail, chow_samples, drop = FALSE])
    )
}

heatmap_df <- bind_rows(pb_heatmap_data)
write_csv(heatmap_df, paste0(tab_path, "mtor_gene_expression_pseudobulk.csv"))

# Heatmap of logFC (HFD vs Chow) per gene per cluster
logfc_mat <- heatmap_df %>%
    dplyr::select(gene, cluster, logFC_HFD) %>%
    pivot_wider(names_from = cluster, values_from = logFC_HFD) %>%
    column_to_rownames("gene") %>%
    as.matrix()

# Drop genes missing from any cluster (filtered by edger_prep in some clusters)
logfc_mat <- logfc_mat[complete.cases(logfc_mat), , drop = FALSE]

# Annotate gene category
gene_cat <- case_when(
    rownames(logfc_mat) %in% mtor_positive ~ "Positive regulator",
    rownames(logfc_mat) %in% mtor_negative ~ "Negative regulator",
    rownames(logfc_mat) %in% mtor_downstream ~ "Downstream target"
)

row_ha <- rowAnnotation(
    Category = gene_cat,
    col = list(Category = c("Positive regulator" = "#D64541",
                             "Negative regulator" = "#3498DB",
                             "Downstream target"  = "#F39C12")),
    annotation_name_gp = gpar(fontsize = 7)
)

col_fun <- colorRamp2(c(-1, 0, 1), c("#3498DB", "white", "#D64541"))

ht <- Heatmap(
    logfc_mat,
    name = "logFC\n(HFD vs Chow)",
    col = col_fun,
    cluster_columns = FALSE,
    cluster_rows = TRUE,
    right_annotation = row_ha,
    row_names_gp = gpar(fontsize = 7),
    column_names_gp = gpar(fontsize = 8),
    column_names_rot = 45,
    border = TRUE,
    column_title = "mTOR pathway genes: HFD vs Chow logFC",
    column_title_gp = gpar(fontsize = 9)
)

pdf(paste0(fig_path, "mtor_gene_heatmap_logfc.pdf"), width = 7, height = 6)
draw(ht)
dev.off()
```

## Panel D — Baseline (Chow) mTOR expression across clusters

```{r panel_d_baseline_heatmap, fig.width=7, fig.height=6}
# Heatmap of absolute logCPM in Chow — do Lepr/Glp1r cells have lower baseline mTOR?
chow_mat <- heatmap_df %>%
    dplyr::select(gene, cluster, logCPM_Chow) %>%
    pivot_wider(names_from = cluster, values_from = logCPM_Chow) %>%
    column_to_rownames("gene") %>%
    as.matrix()

# Drop genes missing from any cluster, then Z-score; drop constant rows
chow_mat <- chow_mat[complete.cases(chow_mat), , drop = FALSE]
chow_z <- t(scale(t(chow_mat)))
chow_z <- chow_z[complete.cases(chow_z), , drop = FALSE]

# Rebuild row annotation for filtered genes
gene_cat_chow <- case_when(
    rownames(chow_z) %in% mtor_positive ~ "Positive regulator",
    rownames(chow_z) %in% mtor_negative ~ "Negative regulator",
    rownames(chow_z) %in% mtor_downstream ~ "Downstream target"
)
row_ha_chow <- rowAnnotation(
    Category = gene_cat_chow,
    col = list(Category = c("Positive regulator" = "#D64541",
                             "Negative regulator" = "#3498DB",
                             "Downstream target"  = "#F39C12")),
    annotation_name_gp = gpar(fontsize = 7)
)

ht_chow <- Heatmap(
    chow_z,
    name = "Z-score\n(across clusters)",
    cluster_columns = FALSE,
    cluster_rows = TRUE,
    right_annotation = row_ha_chow,
    row_names_gp = gpar(fontsize = 7),
    column_names_gp = gpar(fontsize = 8),
    column_names_rot = 45,
    border = TRUE,
    column_title = "mTOR pathway genes: Chow baseline (Z-scored)",
    column_title_gp = gpar(fontsize = 9)
)

pdf(paste0(fig_path, "mtor_gene_heatmap_chow_baseline.pdf"), width = 7, height = 6)
draw(ht_chow)
dev.off()
```

## Panel E — DIO-specific mTOR: Cluster 12 vs POMC direct comparison

Per Tan et al. (2025), mTOR drives leptin resistance specifically in POMC neurons.
We test whether Lepr/Glp1r neurons have lower mTOR activity than POMC in DIO.

```{r panel_e_cluster12_vs_pomc}
mtor_c12_vs_pomc <- arc_diet_labeled[[]] %>%
    filter(seurat_clusters %in% c("12", "29"),
           treatment %in% c("Chow", "HFD")) %>%
    mutate(ct_label  = all_lepr_labels[as.character(seurat_clusters)],
           treatment = factor(treatment, c("Chow", "HFD"))) %>%
    group_by(hash.mcl.ID, hash_pool, treatment, ct_label) %>%
    summarise(mean_mtor_pos  = mean(mTOR_pos_1),
              mean_mtor_neg  = mean(mTOR_neg_1),
              mean_mtor_net  = mean(mTOR_net),
              .groups = "drop")

# Interaction model: cluster × diet with random intercept for animal
# Animals contribute pseudobulk to both cluster 12 and 29, so observations
# from the same animal are not independent
mod_interaction <- lme4::lmer(
    mean_mtor_pos ~ ct_label * treatment + (1 | hash.mcl.ID),
    data = mtor_c12_vs_pomc
)
cat("\n=== Lepr/Glp1r vs POMC × Diet interaction (mTOR positive regulators) ===\n")
print(summary(mod_interaction)$coefficients)
cat("\nType II Anova:\n")
print(car::Anova(mod_interaction, type = "II", test.statistic = "F"))

# Same for net activity
mod_net_int <- lme4::lmer(
    mean_mtor_net ~ ct_label * treatment + (1 | hash.mcl.ID),
    data = mtor_c12_vs_pomc
)
cat("\n=== Lepr/Glp1r vs POMC × Diet interaction (net mTOR activity) ===\n")
print(summary(mod_net_int)$coefficients)
cat("\nType II Anova:\n")
print(car::Anova(mod_net_int, type = "II", test.statistic = "F"))

# Pairwise: Lepr/Glp1r vs POMC within each diet condition
emm_ct <- emmeans(mod_interaction, pairwise ~ ct_label | treatment)
cat("\n=== Pairwise contrasts: Lepr/Glp1r vs POMC within each diet ===\n")
print(summary(emm_ct$contrasts))

write_csv(
    bind_rows(
        as.data.frame(summary(emm_ct$contrasts)) %>% mutate(score = "mTOR_positive"),
        {
            emm_net <- emmeans(mod_net_int, pairwise ~ ct_label | treatment)
            as.data.frame(summary(emm_net$contrasts)) %>% mutate(score = "mTOR_net")
        }
    ),
    paste0(tab_path, "mtor_cluster12_vs_pomc_stats.csv")
)
```

## Panel F — Key individual genes (Mtor, Rptor, Tsc1, Tsc2, Ddit4, Pten)

```{r panel_f_individual_genes, fig.width=10, fig.height=5}
focal_genes <- c("Mtor", "Rptor", "Tsc1", "Tsc2", "Ddit4", "Pten")
focal_genes <- intersect(focal_genes, rownames(arc_diet_labeled))

ind_df <- FetchData(
    arc_diet_labeled,
    vars = c(focal_genes, "seurat_clusters", "treatment", "hash.mcl.ID", "hash_pool")
) %>%
    filter(seurat_clusters %in% all_lepr_clusters,
           treatment %in% c("Chow", "HFD")) %>%
    mutate(ct_label  = all_lepr_labels[as.character(seurat_clusters)],
           ct_label  = factor(ct_label, c("AgRP", "Lepr/Glp1r", "POMC", "Tbx19", "Irx3")),
           treatment = factor(treatment, c("Chow", "HFD"))) %>%
    pivot_longer(cols = all_of(focal_genes), names_to = "gene", values_to = "expr") %>%
    group_by(hash.mcl.ID, hash_pool, treatment, ct_label, gene) %>%
    summarise(mean_expr = mean(expr), .groups = "drop")

p_ind <- ggplot(ind_df, aes(treatment, mean_expr, fill = treatment)) +
    geom_boxplot(outlier.shape = NA, width = 0.6, linewidth = 0.3) +
    geom_jitter(width = 0.15, size = 0.5, alpha = 0.5, shape = 16) +
    facet_grid(gene ~ ct_label, scales = "free_y") +
    scale_fill_manual(values = diet_cols) +
    cowplot::theme_cowplot(font_size = BASE_SIZE) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none",
          strip.background = element_blank()) +
    labs(x = NULL, y = "Mean Expression")

ggsave(paste0(fig_path, "mtor_individual_genes_chow_hfd.pdf"),
       plot = p_ind, h = 5, w = 10)
```

## Summary statistics table

```{r summary_table}
cat("\n========================================\n")
cat("SUMMARY: mTOR pathway in Lepr/Glp1r neurons\n")
cat("========================================\n\n")
cat("Gene sets used:\n")
cat("  Positive regulators:", paste(mtor_positive_present, collapse = ", "), "\n")
cat("  Negative regulators:", paste(mtor_negative_present, collapse = ", "), "\n")
cat("  Downstream targets:", paste(mtor_downstream_present, collapse = ", "), "\n\n")

cat("Key results (Chow vs HFD):\n")
mtor_all_stats %>%
    filter(score == "mTOR_positive_regulators") %>%
    mutate(sig = sig_label(padj),
           direction = ifelse(estimate > 0, "higher in Chow", "higher in HFD")) %>%
    dplyr::select(ct_label, estimate, padj, sig, direction) %>%
    print()

cat("\nInterpretation relative to Tan et al. (2025) hypothesis:\n")
cat("If Lepr/Glp1r neurons show lower mTOR activity increase in HFD compared\n")
cat("to other clusters (especially POMC), this supports the reviewer's suggestion\n")
cat("that relatively lower mTOR activity preserves leptin sensitivity.\n")
```
