---
title: "Supplemental & Extended Data Tables"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup_libraries}
library(tidyverse)

tab_dir <- here::here("output_figures/paper/extended_figures/tables/")
dir.create(tab_dir, recursive = TRUE, showWarnings = FALSE)
```

## Extended Data Table 1 — Xenium 300-gene panel
```{r ed_table1}
# Panel gene list — from the spatial analysis
cat("ED Table 1: Xenium gene panel. Source: sc_preprocessing/xenium_qupath/ panel design files.\n")
cat("This table is assembled from the 10X panel CSV — not generated by code.\n")
```

## Extended Data Table 2 — Pearson correlations (fasting vs DIO) per cluster
```{r ed_table2}
cor_results <- read.csv(here::here("output_figures/paper/extended_figures/overlaps/pearson_fast_hfd_all_clusters.csv"))
write_csv(cor_results, file.path(tab_dir, "ed_table2_pearson_fast_hfd.csv"))
cat("ED Table 2:", nrow(cor_results), "clusters\n")
print(cor_results %>% arrange(r))
```

## Extended Data Table 3 — Gene cluster assignments (963 diet-regulated genes)
```{r ed_table3}
hmap_genes <- read.csv(here::here("output_figures/paper/figure2/tables/glp1r_lepr_hmap_genes.csv"))
write_csv(hmap_genes, file.path(tab_dir, "ed_table3_diet_gene_clusters.csv"))
cat("ED Table 3:", nrow(hmap_genes), "genes\n")
```

## Extended Data Table 4 — LGS genes with loadings
```{r ed_table4}
lgs <- read.csv(here::here("output_figures/paper/lgs/lgs_genes.csv"))
go <- read.csv(here::here("output_figures/paper/lgs/goterm_enrich_lgs.csv"))
write_csv(lgs, file.path(tab_dir, "ed_table4_lgs_genes.csv"))
write_csv(go, file.path(tab_dir, "ed_table4_lgs_go_enrichment.csv"))
cat("ED Table 4:", nrow(lgs), "LGS genes,", nrow(go), "GO terms\n")
```

## Extended Data Table 5 — LGS statistics across populations and conditions
```{r ed_table5}
# Diet LGS
lgs_diet <- read.csv(here::here("output_figures/paper/figure2/tables/lgs_focal_diet_stats.csv"))
# Leptin IP LGS timecourse
lgs_lepip <- read.csv(here::here("output_figures/paper/figure2/tables/lgs_lepip_timecourse_stats.csv"))
# KO LGS
lgs_ko <- read.csv(here::here("output_figures/paper/figure4/tables/lgs_glp1rko_stats.csv"))

write_csv(lgs_diet, file.path(tab_dir, "ed_table5_lgs_diet_stats.csv"))
write_csv(lgs_lepip, file.path(tab_dir, "ed_table5_lgs_lepip_stats.csv"))
write_csv(lgs_ko, file.path(tab_dir, "ed_table5_lgs_ko_stats.csv"))
cat("ED Table 5: diet", nrow(lgs_diet), "rows, lepip", nrow(lgs_lepip), "rows, ko", nrow(lgs_ko), "rows\n")
```

## Extended Data Table 6 — Leptin IP DE genes (Lepr/Glp1r and Agrp)
```{r ed_table6}
glp_de <- read.csv(here::here("output_figures/paper/figure2/tables/glp1r_leptin_degs_wt.csv"))
agrp_de <- read.csv(here::here("output_figures/paper/figure2/tables/agrp_leptin_degs_wt.csv"))
write_csv(glp_de, file.path(tab_dir, "ed_table6_lepip_glp1r_degs.csv"))
write_csv(agrp_de, file.path(tab_dir, "ed_table6_lepip_agrp_degs.csv"))
cat("ED Table 6: Glp1r", sum(glp_de$adj.P.Val < 0.05), "sig genes, Agrp", sum(agrp_de$adj.P.Val < 0.05), "sig genes\n")
```

## Extended Data Table 7 — KO DE genes (Lepr/Glp1r and Agrp, Chow and DIO)
```{r ed_table7}
files <- list(
    glp1r_chow = "glp1rko_clus12_chow_degs.csv",
    glp1r_dio = "glp1rko_clus12_dio_degs.csv",
    agrp_chow = "glp1rko_agrp_chow_degs.csv",
    agrp_dio = "glp1rko_agrp_dio_degs.csv"
)

for (nm in names(files)) {
    d <- read.csv(here::here("output_figures/paper/figure4/tables", files[[nm]]))
    write_csv(d, file.path(tab_dir, paste0("ed_table7_", nm, ".csv")))
    cat(nm, ":", sum(d$adj.P.Val < 0.05), "DEGs (FDR<0.05)\n")
}
```

## Extended Data Table 8 — scDist results (all comparisons, all clusters)
```{r ed_table8}
scdist <- read.csv(here::here("output_figures/paper/extended_figures/overlaps/scdist_all_results.csv"))
write_csv(scdist, file.path(tab_dir, "ed_table8_scdist_all.csv"))
cat("ED Table 8:", nrow(scdist), "cluster-comparison pairs\n")
```

## Extended Data Table 9 — Cluster markers (diet dataset)
```{r ed_table9}
markers <- read.csv(here::here("output_figures/paper/extended_figures/markers/diet_all_markers.csv"))
write_csv(markers, file.path(tab_dir, "ed_table9_cluster_markers.csv"))
cat("ED Table 9:", nrow(markers), "marker genes across", n_distinct(markers$cluster), "clusters\n")
```

## Supplementary Table 1 — LGS genes (same as ED Table 4, for supplementary)
```{r supp_table1}
cat("Supplementary Table 1 = ED Table 4 (LGS genes). Already generated.\n")
```

## Summary
```{r summary}
cat("\n=== All tables generated in:", tab_dir, "===\n")
cat(length(list.files(tab_dir, pattern = "*.csv")), "CSV files\n")
```
